\newpage
\section{Relay configuration}
\label{relay-conf}
Relay configuration is stored in \verb+relay.conf+ file in the
\verb+/etc/dibbler/+ directory (Linux systems) or in current directory
(Windows systems).

\subsection{Global scope}

Every option can be declared in global scope.
Config file consists of global options and one or more inteface
definitions. Note that reasonable minimum is 2 interfaces, as defining
only one would mean to resend messages on the same interface.

\subsection{Interface declaration}

Interface can be declared this way:
\begin{lstlisting}
iface interface-name
{
  interface options
}
\end{lstlisting}

or

\begin{lstlisting}
iface number
{
  interface options
}
\end{lstlisting}

where name\_of\_the\_interface denotes name of the interface and
number denotes it's number. It does not need to be enclosed in
single or double quotes (except windows cases, when interface name
contains spaces).

\subsection{Options}

Every option has a scope it can be used in, default value and
sometimes allowed range.

\begin{description}
 \item[log-level] -- (scope: global, type: integer, default: 7) Defines
            verbose level of the log messages. The valid range if
            from 1 (Emergency) to 8 (Debug). The higher the logging
            level is set, the more messages dibbler will print.
 \item[log-name] -- (scope: global, type: string, default: Client). Defines
            name, which should be used during logging.
 \item[log-mode] -- (scope: global, type: short, full or precise,
            default value: full) Defines logging mode. In the
            default, full mode, name, date and time in the h:m:s format
            will be printed. In short mode, only minutes and
            seconds will be printed (this mode is useful on
            terminals with limited width). Recently added precise
            mode logs information with seconds and microsecond
            precision. It is a useful for finding bottlenecks in
            the DHCPv6 autoconfiguration process.
\item[interface-id-order] -- (scope: global, type: before, after or omit,
        default: before) Defines placement of the
        interface-id option. Options can be placed in the \msg{RELAY-FORW}
        message is arbitrary order. This option has been specified to control
        that order. \opt{interface-id} option can be placed before or after
        \opt{relay-message} option. There is also possibility to instruct
        server to omit the \opt{interface-id} option altogether, but since
        this violates \cite{rfc3315}, it should not be used. In general, this
        configuration parameter is only useful when dealing with buggy relays,
        which can't handle all option orders properly. Consider this parameter
        a debugging feature. Note: similar parameter is available in the dibbler-server.

\item[client multicast] -- (scope: interface, type: boolean, default: false)
        This command instructs dibbler-relay to listen on this particular interface
        for client messages sent to multicast (ff02::1:2) address.
\item[client unicast] -- (scope: interface, type: address, default: not defined)
        This command instructs dibbler-relay to listen to messages sent to a specific
        unicast address. This feature is usually used to connect multiple relays
        together.
\item[server multicast] -- (scope: interface, type: boolean, default: false)
        This command instructs dibbler-relay to send messages (received on any interface)
        to the server multicast (ff05::1:3) address. Note that this is not the same
        multicast address as the server usually listens to (ff02::1:2). Server must be
        specifically configured to be able to receive relayed messages.
\item[server unicast] -- (scope: interface, type: address, default: none)
        This command instructs dibbler-relay to send message (received on any interface)
        to speficied unicast address. Server must be properly configured to to be able to
        receive unicast traffic. See \emph{unicast} command in the \ref{example-server-unicast}
        section.
\item[interface-id] -- (scope: interface, type: integer, default: none)
        This specifies identifier of a particular interface. It is used to generate
        \opt{interface-id} option, when relaying message to the server. This option
        is then used by the server to detect, which interface the message originates from.
        It is essential to have consistent interface-id defined on the relay side and
        server side. It is worth mentioning that interface-id should be specified on the
        interface, which is used to receive messages from the clients, not the
        one used to forward packets to server.
\item[guess-mode] -- (scope: global, type: boolean, default: no)
        Switches relay into so called guess-mode. Under normal operation, client sends
        messages, which are encapsulated and sent to the server. During this encapsulation
        relay appends \opt{interface-id} option and expects that server will use the same
        \opt{interface-id} option in its replies. Relay then uses those \opt{interface-id}
        values to detect, which the original request came from and sends reply to the same
        interface. Unfortunately, some servers does not sent \opt{interface-id} option.
        Normally in such case, dibbler-relay drops such server messages as there is no
        easy way to determine where such messages should be relayed to. However, when
        guess-mode is enabled, dibbler-relay tries to guess the destination interface.
        Luckily, it is often trivial to guess as there are usually 2 interfaces: one
        connected to server and second connected to the clients.
\item[option remote-id] -- (scope: global, type: option, default: none)
        Tells the relay agent to insert remote-id option. It is followed by a
        number (enterprise-id), a dash (``-'') and a hex string that specifies
        the actual content of the remote-id option being
        inserted. Remote-id is specified in \cite{rfc4649}.
\item[option relay-id] -- (scope: global, type: option, default: none)
        Tells the relay agent to insert relay-id option. It takes one parameter,
        which specifies the actual content of the relay-id. The parameter must
        be specified as a hex string.
\item[option link-layer] -- (scope: global, type: option, default: none)
        Tells the relay agent to insert client link-layer address option, as
        specified in \cite{rfc6939}. It should be noted that the source MAC
        address is extracted from incoming link-local (fe80:...) address or
        from client-id. Both methods are not reliable and susceptible for
        spoofing. This was implemented mostly as a testing feature for the
        server implementation.
\end{description}

\subsection{Relay configuration examples}
\label{example-relay}

Relay configuration file is fairly simple. Relay forwards DHCPv6
messages between interfaces. Messages from client are encapsulated and
forwarded as RELAY\_FORW messages. Replies from server are received as
RELAY\_REPL message. After decapsulation, they are being sent back to
clients.

It is vital to inform server, where this relayed message was
received. DHCPv6 does this using interface-id option. This identifier
must be unique. Otherwise relays will get confused when they will
receive reply from server. Note that this id does not need to be
alligned with system interface id (ifindex). Think about it as
"ethernet segment identifier" if you are using Ethernet network or as
"bss identifier" if you are using 802.11 network.

If you are interested in additional examples, download source version
and look at \verb+*.conf+ files.

\subsubsection{Example 1: Simple}
\label{example-relay-1}
Let's assume this case: relay has 2 interfaces: eth0 and
eth1. Clients are located on the eth1 network. Relay should receive
data on that interface using well-known ALL\_DHCP\_RELAYS\_AND\_SERVER
multicast address (ff02::1:2). Note that all clients use multicast
addresses by default. Packets received on the eth1 should be
forwarded on the eth0 interface, using multicast address. See section
\ref{example-server-relay1} for corresponding server configuration.

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short
iface eth0 {
  server multicast yes
}
iface eth1 {
  client multicast yes
  interface-id 5020
}
\end{lstlisting}

\subsubsection{Example 2: Unicast/multicast}
It is possible to use unicast addresses instead/besides of default
multicast addresses. Following example allows message reception from
clients on the 2000::123 address. It is also possible to instruct
relay to send encapulated messages to the server using unicast
addresess. This feature is configured in the next section
(\ref{example-relay-multiple}).

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short
iface eth0 {
  server multicast yes
}
iface eth1 {
  client multicast yes
  client unicast 2000::123
  interface-id 5020
}
\end{lstlisting}

\subsubsection{Example 3: Multiple interfaces}
\label{example-relay-multiple}
Here is another example. This time messages should be forwarded from
eth1 and eth3 to the eth0 interface (using multicast) and to the eth2
interface (using server's global address 2000::546). Also clients must
use multicasts (the default approach):

\begin{lstlisting}
# relay.conf
iface eth0 {
  server multicast yes
}
iface eth2 {
  server unicast 2000::456
}
iface eth1 {
  client multicast yes
  interface-id 1000
}
iface eth3 {
  client multicast yes
  interface-id 1001
}
\end{lstlisting}

\subsubsection{Example 4: 2 relays}
\label{example-relay-cascade}
Those two configuration files correspond to the ,,2 relays'' example
provided in section \ref{example-server-relay2}. See section
\ref{feature-relays} for detailed exmplanations.

\begin{lstlisting}
# relay.conf - relay 1
log-level 8
log-mode full

# messages will be forwarded on this interface using multicast
iface eth2 {
   server multicast yes    // relay messages on this interface to ff05::1:3
 # server unicast 6000::10 // relay messages on this interface to this global address
}

iface eth1 {
#  client multicast yes    // bind ff02::1:2
  client unicast 6011::1   // bind this address
  interface-id 6011
}
\end{lstlisting}

\begin{lstlisting}
# relay.conf - relay 2
iface eth0 {
#   server multicast yes  // relay messages on this interface to ff05::1:3
  server unicast 6011::1  // relay messages on this interface to this global address
}

# client can send messages to multicast
# (or specific link-local addr) on this link
iface eth1 {
  client multicast yes    // bind ff02::1:2
# client unicast 6021::1  // bind this address
  interface-id 6021
}
\end{lstlisting}

\subsubsection{Example 5: Guess-mode}
In the 0.6.0 release, a new feature called guess-mode has been
added. When client sends some data and relay forwards it to the
server, it always adds interface-id option to specify, which link
the data has been originally received on. Server, when responding to
such request, should include the same interface-id option in the
reply. However, in some poor implementations, server fails to do
that. When relay receives such poorly formed response from the server,
it can't decide which interface should be used to relay this
message.

Normally such packets are dropped. However, it is possible to switch
relay into a guess-mode. It tries to find any suitable interface,
which it can forward data on. It is not very reliable, but sometimes
it is better than dropping the message altogether.

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short
guess-mode

iface eth0 {
  server multicast yes
}
iface eth1 {
  client multicast yes
  interface-id 5020
}
\end{lstlisting}

\subsubsection{Example 6: Relaying to multicast}
During normal operation, relay sends forwarded messages to a
\emph{All\_DHCP\_Servers} (FF05::1:3) multicast address.

Although author does not consider this an elegant solution, it is also
possible to instruct relay to forward message to a \emph{All\_DHCP\_Relay\_Agents\_and\_Servers}
(ff02::1:2) multicast address. That is quite convenient when there are several
relays connected in a cascade way (server -- relay1 -- relay2 -- clients).

For details regarding DHCPv6-related multicast addresses and relay operation, see \cite{rfc3315}.

To achieve this behavior, \emph{server unicast} can be used. Note that
name of such parameter is a bit misleading (``server unicast'' used to specify
multicast address). That parameter should be rather called ``destination address'',
but to maintain backward compatibility, it has its current name.

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short

iface eth0 {
  server unicast ff02::1:2
}
iface eth1 {
  client multicast yes
  interface-id 5020
}
\end{lstlisting}

\subsubsection{Example 7: Options inserted by the relay}
Typically relay agent receives messages from clients, encapsulates
them and sends towards the server. The only option being added is
interface-id option (if specified). However, in some cases it makes
sense for the relay agent to insert additional options. Dibbler relay
supports several options that can be inserted:

\begin{description}
\item[remote-id] -- Remote-id option may identify the remote client.
Dibbler implementation is very simple, as it allows setting only
one specific value, not unique value, one for each client. That makes
this feature useful for testing purposes, but its deployment in a
production network is unlikely. Remote-id takes two parameters. The
first one is enterprise-id, a 32bit unique identifier that
characterises a vendor. The second parameter is arbitrary length hex
string. Remote-id is specified in \cite{rfc4649}.
\item[echo-request] -- In some cases, relay inserts options and would
like the server to send them back in its responses. Typically, those
options are then processed by the relay to correctly send back to the
client. Dibbler relay does not need such options to operate, but is
able to simulate them. It is possible to specify one or more options
that the relay requests the server to echo back. This option takes a
list of coma separated values that designate option codes. Dibbler
relay will insert options with specified option codes, but they will
not carry any useful value. Echo Request Option is specified
in \cite{rfc4994}.
\item[relay-id] -- Relay agent may be configured to insert
interface-id option. However, that option identifies an interface
within each relay, but not necessarily the relay itself. In some cases
it is useful for the relay agent to insert an option that identifies
itself. That is implemented as relay-id option. This is particularly
useful for bulk leasequery. This option is defined in \cite{rfc5460}.
\item[link-layer] -- The initial DHCPv6 spec lacked information about client
MAC addresses. In principle, the server is able to discover client's
MAC address when receiving direct traffic. Unfortunately, it can't do
that if the packet traverses a relay agent. That deficiency was
addressed in \cite{rfc6939}. It defines an option that relays can
insert when receiving incoming traffic from a client. Note that this
implementation is not perfect. Instead of getting that information
from layer 2 directly (there is no API to do that as of late 2014), it
tries to get the information from client-id (DUID-LLT or DUID-LL) or
from message source address (if it is link-local and uses
EUI-64). This is not a bullet-proof solution, but it should work in
most cases.
\end{description}

\begin{lstlisting}
log-level 8
log-mode short

# Uncomment the following line to force relay to start including remote-id
# option with entreprise-id 5 and content of 01:02:03:04
option remote-id 5-0x01020304

# Uncomment the following line to force relay to send Echo Request Option
# asking the server to echo back options 100, 101 and 102
option echo-request 100,101,102

# Uncommenting this option will make relay to insert relay-id option
# into forwarded RELAY-FORW messages.
option relay-id aa:bb:cc:dd:ee:ff

# Uncomment this line to tell relay to attempt to insert client link-layer
# address option. Relay will attempt to get that info from client's DUID
# or source IPv6 link-layer address. These are not 100% reliable methods!
option link-layer

# The relay should listen on eth1 interface for incoming client's traffic.
# Clients by default send their traffic to multicast.
iface eth1 {
  client multicast yes

  # When forwarding traffic from that interface, please add interface-id
  # with value 5555, so the server will know where the clients are connected.
  interface-id 5555
}

# This is a second interface. It is used to reach the server.
iface eth2 {

  # Send message on this interface to the server multicast (ff05::1:3)
  server multicast yes
}
\end{lstlisting}

\newpage

\section{Requestor configuration}
Requestor (entity used for leasequery) does not use configuration
files. All parameters are specified by command-line switches. See
section \ref{feature-leasequery} for details.
