%%
%% Dibbler - a portable DHCPv6
%%
%% authors: Tomasz Mrugalski <thomson@klub.com.pl>
%%          Marek Senderski <msend@o2.pl>
%%
%% released under GNU GPL v2 or later licence
%%
%% $Id: dibbler-devel-arch-clnt.tex,v 1.3 2004-11-25 01:16:36 thomson Exp $
%%
%% $Log: not supported by cvs2svn $
%% Revision 1.2  2004/10/25 20:45:54  thomson
%% Option support, parsers rewritten. ClntIfaceMgr now handles options.
%%
%% Revision 1.1  2004/07/05 01:04:40  thomson
%% Initial version.
%%
%%

\section{Common Architecture}

General architecture is common between server and client. In both
cases, there are 4 managers:
\begin{itemize}
\item[IfaceMgr] -- Interface Manager. It represents all network interfaces present in the
  system. They're represented by TIfaceIface objects and stored in
  IfaceLst. Each interface has list of open sockets, represented with
  TIfaceSocket objects. There are also a number of auxiliary functions
  for getting proper interface. 
\item [AddrMgr] -- Address Manager. It is an address database, which
  stores all informations about clients, IAs and associated addresses.
\item [CfgMgr] -- Config Manager. It is being used to read
  configuration information from config file and provide those
  informations while runtime.
\item [TransMgr] -- Transmission Manager, sometimes called Transaction
  Manager. It is responsible for network interaction and logic. It
  sends various messages when such need arise, matches received
  responses with sent messages, retransmits etc. It contains list of
  messages currently being used.
\end{itemize}

They are all created within DHCPClient or DHCPServer object. 

\section{Client Architecture}

Client is represented by a DHCPClient object. It contains 4 large
managers, each with its own functions:
\begin{itemize}
\item IfaceMgr
\item AddrMgr
\item CfgMgr
\item TransMgr

\end{itemize}

