\newpage

\newcommand{\rfc}[1]{\href{http://tools.ietf.org/html/rfc#1}{RFC#1}}
\begin{thebibliography}{1}

\bibitem{rfc2030} Mills, D., ``Simple Network Time Protocol (SNTP) Version 4 for
        IPv4, IPv6 and OSI'', \rfc{2030}, IETF, October 1996.

\bibitem{rfc3263} J.Rosenberg and H. Schulzrinne, ``Session Initiation Protocol
       (SIP): Locating SIP Servers'', \rfc{3263}, IETF, June 2002.

\bibitem{rfc3315} R. Droms, Ed. ``Dynamic Host Configuration Protocol
  for IPv6 (DHCPv6)'', \rfc{3315}, IETF, July 2003.

\bibitem{rfc3319} H. Schulzrinne, and B. Volz ``Dynamic Host
  Configuration Protocol (DHCPv6) Options for Session Initiation
  Protocol (SIP) Servers'', \rfc{3319}, IETF, July 2003.

\bibitem{rfc3596} S. Thomson, C. Huitema, V. Ksinant and M. Souissi
  ``DNS Extensions to Support IP Version 6'', \rfc{3596}, IETF, October
  2003.

\bibitem{rfc3633} O. Troan, and R. Droms ``IPv6 Prefix Options for
  Dynamic Host Configuration Protocol (DHCP) version 6'', \rfc{3633},
  IETF, December 2003.

\bibitem{rfc3646} R. Droms, Ed. ``DNS Configuration options for
  Dynamic Host Configuration Protocol for IPv6 (DHCPv6)'', \rfc{3646},
  IETF, December 2003.

\bibitem{rfc3736} R. Droms, ``Stateless Dynamic Host Configuration
  Protocol (DHCP) Service for IPv6'', \rfc{3736}, IETF, April 2004.

\bibitem{rfc3898} V. Kalusivalingam ``Network Information Service
  (NIS) Configuration Options for Dynamic Host Configuration Protocol
  for IPv6 (DHCPv6)'', \rfc{3898}, IETF, October 2004.

\bibitem{rfc4033} R. Arends,  R. Austein, M. Larson, D. Massey and S. Rose
  ``DNS Security Introduction and Requirements'',
  \rfc{4033}, IETF, March 2005

\bibitem{rfc4242} S. Venaas, T. Chown, and B. Volz
 ``Information Refresh Time Option for DHCPv6'', \rfc{4242},
 IETF, Nov. 2005.

\bibitem{rfc4075} V. Kalusivalingam ``Simple Network Time Protocol
        (SNTP) Configuration Option for DHCPv6'', \rfc{4075}, IETF, May 2005.

\bibitem{rfc4704} M. Stapp and  B.Volz ``The Dynamic Host
Configuration Protocol for IPv6 (DHCPv6) Client Fully Qualified Domain
Name (FQDN) Option'', \rfc{4704}, IETF, October 2006

\bibitem{rfc4861} T.Narten, E.Nordmark, W.Simpson and H.Soliman,
``Neighbor Discovery for IP version 6 (IPv6)'', \rfc{4861}, IETF,
September 2007

\bibitem{rfc4862} S.Thomson, T.Narten, T.Jinmei, ``IPv6 Stateless
Address Autoconfiguration'', \rfc{4862}, IETF, September 2007

\bibitem{rfc5007} J. Brzozowski, K. Kinnear, B. Volz and S. Zeng
  ``DHCPv6 Leasequery'', \rfc{5007}, IETF, September 2007

\bibitem{rfc6333}  A.Durand, R.Droms, J.Woodyatt, Y.Lee,
``Dual-Stack Lite Broadband Deployments Following IPv4 Exhaustion'',
\rfc{6333}, IETF, Aug. 2011

\bibitem{rfc6334} D.Hankins, T.Mrugalski, ``Dynamic Host
Configuration Protocol for IPv6 (DHCPv6) Options for Dual-Stack
Lite'', \rfc{6334}, IETF, Aug. 2011

\bibitem{draft-aaa} Vishnu Ram, Saumya Upadhyaya, Nitin Jain
        ``Authentication, Authorization and key management for
        DHCPv6'', work in progress (expired),
        draft-ram-dhc-dhcpv6-aakey-01, IETF, August 2006

\bibitem{phd} T. Mrugalski, ``Optimization of the autoconfiguration
mechanisms of the mobile stations supporting IPv6 protocol in the IEEE
802.16 environmentdfdf'', Ph.D dissertation, Gdañsk, Oct. 2009

\bibitem{networks2010} T. Mrugalski, J.Wozniak, K.Nowicki, ``Remote DHCPv6
Autoconfiguration for Mobile IPv6 nodes'', IEEE 14th International
Telecommunications Network Strategy and Planning Symposium, Warsaw,
Poland, Sept. 2010

\bibitem{atnac2010} T.Mrugalski, J.Wozniak, K.Nowicki, ``Remote
Stateful Autoconfiguration for Mobile IPv6 Nodes with Server Side
Duplicate Address Detection'', IEEE, Australasin Telecommunication
Networks and Applications Conference, Auckland, New Zealand, Nov. 2010

\bibitem{draft-route-option} W.Dec, T.Mrugalski, T.Sun, B. Sarikaya,
``DHCPv6 Route Options'', MIF WG, work in progress,
draft-ietf-mif-dhcpv6-route-option-03, IETF, Sep. 2011
.

\bibitem{draft-remote-autoconf} T.Mrugalski, ``Remote DHCPv6
Autoconfiguration'', work-in-progress (expired), IETF, July 2010

\bibitem{draft-timezone} A.K. Vijayabhaskar ``Time Configuration Options
        for DHCPv6'', work in progress (expired),
        draft-ietf-dhc-dhcpv6-opt-timeconfig-03, IETF, October 2003

\end{thebibliography}
\addcontentsline{toc}{section}{Bibliography}%\hfill \thepage\\}
