%%
%% Dibbler - a portable DHCPv6
%%
%% authors: Tomasz Mrugalski <thomson@klub.com.pl>
%%          Michal Kowalczuk <michal@kowalczuk.eu>
%%
%% released under GNU GPL v2 licence
%%

\newpage
\section{Configuration files}

This section describes Dibbler server, relay  and client
configuration. Square brackets denotes optional values: mandatory
[optional]. Alternative is marked as $\mid$. A $\mid$ B means A or
B. Parsers are case-insensitive, so Iface, IfAcE, iface and IFACE mean
the same. This does not apply to interface names. eth0 and
ETH0 are dwo diffrent interfaces.

\subsection{Data types}
Config file parsing is token-based. Token can be considered a keyword
or a specific phrase. Here are more commonly used types:
\begin{description}
\item[IPv6 address] -- IPv6 address, e.g. 2000:db8:1::dead:beef
\item[32-bit decimal integer] -- string containing only numbers, e.g. 12345
\item[string] -- string of arbitrary characters enclosed in single or double
  quotes, e.g. 'this is a string'. If string contains only a-z, A-Z and
  0-9 characters, quotes can be omited, e.g. beeblebrox
\item[DUID identifier] -- hex number starting with 0x,
  e.g. 0x12abcd. In Dibbler version 0.8.0RC1, another format was
  introduced: 2 hex digits separated by comma, e.g. 12:aa:bb:cc:d5. As
  this format may in some cases be confused with IPv6 address, the old
  format (starting with 0x) remains to be supported.
\item[IPv6 address list] -- IPv6 addresses separated with commas,
           e.g. 2001:db8:1::face:b00c, fe80::abcd:1234,::1
\item[DUID list] -- DUIDs separated with commas, e.g. 0x0123456,0x0789abcd
\item[string list] -- strings separated with comas, e.g. tealc,jackson,carter,oneill
\item[boolean] -- YES, NO, TRUE, FALSE, 0 or 1. Each of them can be
  used, when user is expected to enable or disable specific option.
\end{description}

\subsection{Scopes}
\label{scope}
There are four scopes, in which options can be specified: global,
inteface, IA and address. Every option is specific for one scope.
Each option is only applied to a scope and all subscopes in which it is
defined. For example, T1 is defined for
IA scope. If there are several interfaces and each has several address
classes, repeating the same T1 value many time may be a bit awkward.
Therefore parameters can be also used in more common scopes. In this
case -- in interface or global. Defining T1 in interface scope means:
,,for this interface the default T1 value is ...''. The same applies
to global scope. Options can be used multiple times. In that case
value defined later is used.

Global scope is the largest. It covers the whole config file and
applies to all intefaces, IAs, and addresses, unless some lower scope
options override it. Next scope is inteface. Options defined there
are inteface-specific and apply to a given interface, all IAs in this
interface and addresses in those IAs. Next is IA scope. Options
defined there are IA-specific and apply to this IA and to addresses it
contains. The narrowest scope is address or prefix.

\subsection{Comments}

Comments are allowed in configuration files. All common comment styles are supported:
\begin{itemize}
\item C++ style one-line comments: \verb+// this is comment+
\item C style multi-line comments: \verb+ /* this is multi-line comment */+
\item script style one-line comments: \verb+# this is one-line comment+
\end{itemize}

\subsection{Client configuration file}
\label{client-cfg-file}
Client configuration file should be named \verb+client.conf+. It
should be placed in the \verb+/etc/dibbler/+ directory (Linux system)
or in the current directory (Windows systems). One of design
requirements for client was ,,out of the box'' usage. To achieve this,
simply use empty \verb+client.conf+ file. Client will try to get one
address for each up and running interface. More preciselt client tries
to configure each up, multicast-capable and running interface, which
has link address at least 6 bytes long. This usually means that client
running in auto-detection mode will not configure tunnels, which
usually have IPv4 address (4 bytes long) as their link address. It
should configure all wired (Ethernet) and wireless (802.11)
interfaces, though.

\subsubsection{Interface declaration}
\label{scope-iface}
Each system interface, which should be configured, must be mentioned in
the configureation file. Interfaces can be declared with this syntax:
\begin{lstlisting}
iface interface-name
{
  interface-options
  IA-options
  address-options
}
\end{lstlisting}

or

\begin{lstlisting}
iface interface-number
{
  interface-options
  IA-options
  address-options
}
\end{lstlisting}

In the latter case, interface-number denotes interface index (or
ifindex). It can be obtained from \verb+ip~l+ (Linux), \verb+ipv6~if+
(Windows) or sometime \verb+ifconfig+ on other systems.
\verb+interface-name+ is an interface name.  Name of the
interface does not have to be enclosed in single or double quotes. It
is necessary only in Windows systems, where interface names sometimes
contain spaces, e.g. ''local network connection''.  Interface scoped
options can be used here. IA-scoped as well as address scoped options
can also be used. They will be treated as a default values for future
definitions of the IA and address instantations.

\subsubsection{IA declaration}
\label{scope-ia}
IA is an acronym for Identity Association. It is a logical entity
representing address or addresses used to perform some functions. It
is not suitable for prefixes (see Section \ref{pd-declaration}).
IA-options can be defined, e.g. T1. IPv6 addresses can be defined
here. All those values will be used as hints for a server.  Almost
always, each DHCPv6 client will have exactly one IA on each
interface. IA is declared using following syntax:

\begin{lstlisting}
ia [iaid]
{
  IA-options
  address-options
  address-declaration
}
\end{lstlisting}

IAID is an optional number, which describes identifier of given IA. If
not specified, it will be automatically assigned integer numbers
starting with 1. There may be more than one IA defined on an
interface. IA and PD (see Section \ref{pd-declaration}) may be used
together.

\subsubsection{TA declaration}
\label{scope-ta}
TA (Temporary Address) is a mechanism very similar to IA that allows
configuration of temporary addresses. See Section \ref{scope-ia}.

\subsubsection{PD declaration}
\label{pd-declaration}
\label{scope-pd}
PD (Prefix Delegation) is a mechanism that allows leasing prefixes in
similar way as addresses. For details, see \cite{rfc3633}. PD in
client config file causes client to send IA\_PD option. This option
informs server that client is requesting prefix delegation.

\begin{lstlisting}
pd [iaid]
{
  pd-options
  prefix-declaration
}
\end{lstlisting}
IAID is an optional number, which identifies this particular PD. If
not specified, it will be automatically assigned integer numbers
starting with 1. There may be more than one PD defined on an
interface. IA (see Section \ref{scope-ia}) may be used together.

\subsubsection{Address declaration}
\label{scope-addr}
When IA is defined, it is sometimes useful to define its address. Its
value will be used as a hint for the server. Address is declared in the
following way:

\begin{lstlisting}
address [number]
{
  address-options
  address-declaration
}
\end{lstlisting}
where number is optional and denotes how many addresses with those values should be
requested. If it is diffrent than 1, then IPv6 address options are not
allowed. Only address scoped options can be used here. Address can be
defined only within IA scope.

\subsubsection{Prefix declaration}
\label{scope-prefix}
When PD is defined, it is sometimes useful to define its prefix. Its
value will be used as a hint for the server. Prefix is declared in the
following way:

\begin{lstlisting}
prefix [number]
{
  prefix-options
  prefix
}
\end{lstlisting}

\subsubsection{Client Configuration Parameters}
\label{client-conf-reference}
Dibbler client supports multiple parameters. Some ore defined by the
base DHCPv6 specification (a RFC 3315 document \cite{rfc3315}),
e.g. IA address option. Other parameters are for definition of one of
multitude of extensions that were defined for DHCPv6 protocol
(e.g. prefix delegation option). Finally, there are many configuration
parameters that are not options in DHCPv6 sense, but rather affect the
way the software operates (e.g. log level). All those parameters may
be defined in client config file. Every statement has defined
scope. See Section~\ref{scope} for details. In many cases, parameters
may also be defined in scopes larger than its default scope. For
example, instead of configuring DNS server option on 3 interfaces, it
can be defined once in global scope.

\begin{description}
\item[iface] -- (scope: global). Takes one parameter that can be
either string (interface name) or integer (interface index). Defines
that client should perform some actions on specified interface. Exact
operations are defined within interface scope. Can optionally take
second string parameter with the only allowed value of
``no-config''. If it is present, it instructs client to not perform
any operation on said interface. See Section \ref{scope-iface}.

\item[ia] -- (scope: interface). Defines IA\_NA (Identity Association for
Non-temporary Addresses), often abbreviated as IA. That is a
container for ``regular'' (non-temporary in DHCPv6 nomenclature)
addresses. Simply saying, this is a client's request for a single normal
address. There may be more than one ia defined on one interface. In
such case, client will request several addresses. It may have one
optional integer parameter that defines unique indentifier (IAID). If
followed by curly brackets, it will create new IA scope. See
Section \ref{scope-ia}.

\item[ta] -- (scope: interface). Defines IA\_TA (Indentity Association
for Temporary Addresses), often abbreviated as TA. This is a
container for temporary addresses. Simply saying, this is a client's
request for a temporary address. If followed by curly brackets, it
will create new IA scope, similar to IA. See
Section \ref{scope-ia}. Note that TA scope accepts only limited set of
parameters (e.g. iaid).

\item[pd] -- (scope: interface). Defines IA\_PD (Identity Association
for Prefix Delegation), often abbreviated as PD. This is a container
for prefixes. Simply saying, this is client's request for prefix
delegation. It may have one optional integer parameter that defines
unique indentifier (IAID). If followed by curly brackets, it will
create new PD scope. See Sections \ref{scope-pd}
and \ref{feature-prefix}.

\item[address] -- (scope: ia or ta). Defines an IPv6 address. It is
usually defined in IA or TA to specify that an address should be sent. It
takes one optional parameter that defines multiplier. For example, to
define 3 addresses, following syntax may be used:
\begin{lstlisting}
address 3 { }
\end{lstlisting}
If followed by curly brackets, creates new address scope. See
Section \ref{scope-addr}.

\item[prefix] -- (scope: pd). Defined an IPv6 prefix. It is defined in
PD to specify that a prefix should be sent. May be empty
(\verb+prefix+) or accompanied with prefix definition that consists of
IPv6 address followed by slash and prefix length (e.g.
\verb+prefix 2001:db8:1::/56+. If followed by curly brackets, creates
new prefix scope. See Section \ref{scope-prefix}.

\item[work-dir] -- (scope: global). Takes one string
 parameter. Defines working directory.

\item[log-level] -- (scope: global). Takes one integer
parameter. Defines verbose level of the log messages. The valid range
is from 1 (very quiet) to 8 (very verbose). Those values are modelled
after levels used in syslog. These are: 1(Emergency), 2(Alert),
3(Critical), 4(Error), 5 (Warning), 6(Notice), 7(Info) and
8(Debug). Currently Dibbler is using levels 3 to 8, as 1 and 2 are
reserved for system wide emergency events.

\item[log-name] -- (scope: global). Takes one string
parameter. Defines than name, which will be used during logging.

\item[log-mode] -- (scope: global). Takes one parameter that can be
   short, full, precise or syslog. Defines logging mode. In the
   default, full mode, name, date and time in the h:m:s format will be
   printed. In short mode, only minutes and seconds will be printed
   (this mode is useful on terminals with limited width). Precise mode
   logs information with seconds and microsecond precision. It is a
   useful as a performance diagnostic tool for finding bottlenecks in
   the DHCPv6 autoconfiguration process. Syslog works under POSIX
   systems (Linux, Mac OS X, BSD family) and allows default POSIX
   logging functions.

\item[log-colors] -- (scope: global). Takes one boolean parameter.
  Defines if logs printed to console should use colors. That feature
  is used to enhance logs readability.  As it makes the log files
  messy on systems that do not support colors, it is disabled by
  default. The default is off.

 \item[strict-rfc-no-routing] -- (scope: global) During normal
  operation, DHCPv6 client should add IPv6 address only
  (i.e. configure it with /128 prefix), without configuring routing.
  Routing is expected to be configured with i.e. router
  advertisements \cite{rfc4861}. However, this behavior is confusing
  and lots of users complained about it. Therefore since the 0.5.0RC1
  release, this has been changed in Dibbler. Right now when Dibbler
  client configures address, it also configures routing, so every host
  is able to communicate with other hosts, which have obtained address
  from the same server. If you don't like this behavior, you might
  want to use this option. Note that draft \cite{draft-route} was
  introduced to solve this DHCPv6 protocol limitation. It is expected
  to be supported in the next Dibbler release.

\item[experimental] -- (scope: global). Allows enabling experimental
features. There are some highly-experimental features present in
Dibbler. To make a clear statement about their experimental nature,
user is required to acknowledge that fact by putting this statement in
its config file. This statement may be present or absent. The default
is absent.

\item[addr-params] -- (scope: IA). Allows configuration of additional
sub-option conveyed in IAADDR. It supplements the usual information
about an address received from a server with prefix length. For
example, if client received address 2001:db8:1::abcd and addr-params
option contains 64, Dibbler client will configure prefix
2001:db8:1::/64 on the interface that was used to communicate with
server. This is experimental feature, not defined in any standard or
draft. Requires \emph{experimental} statement. See
Section \ref{feature-addr-params}.

\item[remote-autoconf] -- (scope: interface). Defines that remote
autoconfiguration should be performed on a given interface. This is
experimental feature, so it requires \emph{experimental}
statement. See Section \ref{feature-remote-autoconf}.

\item[ddns-protocol] -- (scope: global). Takes one string
parameter. Defines protocol that should be used during DNS Update
mechanism. Allowed values are \verb+tcp+, \verb+udp+ and \verb+any+.
Any means that UDP will be tried first and if it fails, update will be
retried over TCP. See Section \ref{feature-dns-update}.

\item[ddns-timeout] -- (scope: global). Takes one integer parameter
that specifies timeout in milliseconds. Defines how long client should
wait for DNS server response during DNS Update before declaring
update a failure. See Section \ref{feature-dns-update}.

\item[script] -- (scope: global). Takes one string parameter that
  specifies script name. When dibbler client receives some options it normally
  sets them up in the system on its own. However, besides of setting up all
  parameters directly, dibbler client can execute external
  script. See Section \ref{feature-client-script} for details.

\item[stateless] -- (scope: global). It may be present or missing. The
  default is missing. Defines that client should run in stateless
  mode. In this mode only configuration parameters are defined, not
  addresses or prefixes. It is mutually exclusive
  with \emph{ia}, \emph{ta} and \emph{pd}. See
  Section \ref{feature-stateless-stateful}. \emph{No-ia}, an alias to that
  command used to be supported, but due to misleading name its support
  was dropped in 0.8.1RC1.


\item[anonymous-inf-request] -- (scope: global). When running in a
  stateless mode, client does not ask for addresses or prefixes, but
  rather requests some general options. By default, it sends its
  client identifier (DUID) to the server. However, it is possible to
  omit this identifier, so the \msg{INF-REQUEST} messages will be
  anonymous. This global option causes client to act in such anonymous
  way.

 \item[inactive-mode] -- (scope: global). This parameter may be
   present or absent. The default is absent. Normally (with
   inactive-mode disabled) client tries to bind all interfaces defined
   in configuration file. If such attempt fails, client reports an
   error and gives up. In some cases it is possible that interface is
   not ready yet, e.g. WLAN interface did not complete association. It
   is possible to modify client behavior, so it will accept downed and
   not running interfaces. To do so, inactive-mode must be enabled. In
   this mode, client will accept inactive interfaces, will add them to
   inactive list and will periodically monitor its state. When the
   interface finally becomes available, client will try to configure
   it. See section \ref{feature-inactive-mode} for details.

 \item[insist-mode] -- (scope: global). Client can be instructed to
   obtain several configuration options, like DNS server configuration
   or domain name. It is possible that server will not provide all
   requested options. Older versions of the dibbler client had been
   very aggressive in such case. It tried very hard to obtain options
   that user specified in config file. To do so, it did
   send \msg{INF-REQUEST} to obtain such option. This behavior has
   changed. Right now when client does not receive all requested
   options, it will complain, but will take no action. To enable old
   behavior, so called insist-mode has been added. Insist-mode will
   also affect the way addresses are requested. If address was
   specified in config file, client will request it in \msg{REQUEST}
   message, rather than sending address offered by server
   in \msg{ADVERTISE} as it is typically done. See
   Section \ref{feature-insist-mode} for details.

 \item[skip-confirm] -- (scope: global). Support for \msg{CONFIRM} messages
   was added in 0.8.0RC1. With it, client may send \msg{CONFIRM} when link
   state change is detected (e.g. switching to possibly new WiFi
   access-point or replugging Ethernet cable). Client also sends
   \msg{CONFIRM} after restart if there are still valid leases found
   in locally stored databased. \emph{skip-confirm} will disable any
   actions that would result in \msg{CONFIRM} transmissions. In
   particular, link state will not be detected and client will ignore
   its previous address during startup.

%%% TODO: Add reconfigure accept here



 \item[duid-type] -- (scope: global). Takes one parameter. Allowed
   values are DUID-LLT, DUID-LL or DUID-EN. The default is DUID-LLT.
   This parameter defines, what type of DUID should be generated if
   there is no DUID already present. If there is a file containing
   DUID, this directive has no effect. DUID-LLT means that DUID will
   be based on link layer address as well as time. DUID-LL means that
   only link layer address will be used. The last value -- DUID-EN --
   Enterprise Number-based generation has a slightly different syntax:
 \begin{lstlisting}
   duid-type duid-en enterprise-number enterprise-id.
\end{lstlisting}
   For example: \verb+duid-type duid-en 1234 0x6789abcd+ means that
   enterprise number is set to 1234 and unique number from that
   company's pool is 67:89:ab:cd (hexadecimal value of arbitrary
   length). See section \ref{feature-duid-types} for details.

\item[option fqdn-s] -- (scope: global). Takes one boolean parameter
  and has the default value of 1. The S bit is used in FQDN option. It
  is used to negotiate, which side (server or client) wants to perform
  DNS Update procedure. See \cite{rfc4704} for details. In general, if
  you don't know that this option does, you don't want to modify this.

\item[option fqdn] -- (scope: interface). Takes optional domain
   name as parameter. This option instructs client to send FQDN
   option. This option has 2 purposes. The first one is to negotiated
   or request Fully Qualified Domain Name for this client. The second
   one is to negotiate, who (client or server) should perform DNS
   Update. If optional parameter is specified, it will be sent in the
   FQDN option. Otherwise FQDN will be sent with empty name. This
   option is defined in \cite{rfc4704}. See
   Section \ref{feature-dns-update} for details.

 \item[rapid-commit] -- (scope: interface). Takes one boolean
   parameter. The default is 0. This option allows rapid commit
   procedure to be performed. Note that enabling rapid-commit on the
   client side is not enough. server must be configured to allow rapid
   commit, too.

 \item[unicast] -- (scope: interface). Takes one boolean
  parameter. The default value is 0. This option specifies if client
  should request unicast communication from the server. If server is
  configured to allow it, it will add unicast option to its
  replies. It will allow client to communicate with server via unicast
  addresses instead of usual multicast.

 \item[preferred-servers] -- (scope: interface). Takes list of
  addresses or list of DUIDs. The default value is empty.  This list
  defines, which servers are preferred. When client sends \msg{SOLICIT}
  message, all servers available in the local network will
  respond. When client receives multiple \msg{ADVERTISE} messages, it
  will choose those sent by servers mentioned on the preferred-server
  list. Otherwise the best server will be chosen. Note that this
  parameter used to be spelled differently (single R). Old spelling
  is still supported, but is deprecated and will be removed soon.

 \item[reject-servers] -- (scope: interface). Takes list of addresses
  or list of DUIDs. The default value is empty. This list defines
  which server must be ignored. It has opposite meaning to the
  prefered-servers list. Servers mentioned here will never be chosen.

 \item[vendor-spec] -- (scope: interface). This option allow
   requesting for a vendor specific configuration option or
   options. Although there are no vendor-specific (i.e. specific for
   Dibbler) parameters, it can be used to test some other DHCPv6
   server implementations. It takes coma separated list following
   tokens: type (integer), type(integer) -- enterprise(number). It
   allows definition of just vendor-specific option and/or
   vendor-specific option with specific enterprise. See feature
   description in Section \ref{feature-vendor-spec}.

 \item[T1] -- (scope: IA or PD) Takes one parameter that specifies T1
   hint sent to a server. The default value is $2^{32}-1$ and is
   expressed in seconds. T1 value assigned by server defines after
   what time client should start renew process. This is only a hint
   for the server. Actual value will be provided by the server.

\item[T2] -- (scope: IA or PD). Takes one parameter that specifies T2 hint
   sent to a server. The default value is default:$2^{32}-1$ and is
   expressed in seconds. This value defines hint for T2. T2 assigned
   by server defined after what time client will start emergency
   rebind procedure if renew process fails. This is only a hint for
   the server. Actual value will be provided by the server.

\item[valid-lifetime] -- (scope:address or prefix). Takes one integer
   parameter that defined requested valid lifetime for address or
   prefix. The default value is $2^{32}-1$. This parameter is
   expressed in seconds. This parameter defines valid lifetime of an
   address. It will be used as a hint for a server, when the client
   sends \msg{REQUEST} messages.

\item[preferred-lifetime] -- (scope:address or prefix).
   type: integer, default:$2^{32}-1$) This parameter defines prefered
   lifetime of an address. It will be used as a hint for a server,
   when there client sends \msg{REQUEST} messages. Note that this
  parameter used to be spelled differently (single R). Old spelling
  is still supported, but is deprecated and will be removed soon.

 \item[option dns-server] -- (scope: interface). Takes optional
   address list as parameter. This option conveys information about
   DNS servers available. After retriving this information, client
   will be able to resolve domain names into IP (both IPv4 and IPv6)
   addresses. If address list is not specified, client will just
   request dns-server option from a server. If it is specified, listed
   addresses will be sent to server as hints. Defined in \cite{rfc3596}.

 \item[option domain] -- (scope: interface). Takes optional domain
   list as parameter. This option is used for retriving domain or
   domains names, which the client is connected in. For example, if
   client's hostname is \verb+alice.mylab.example.com+ and it wants to
   contact \verb+bob.mylab.example.com+ it can simply refer to it
   as \verb+bob+. Without domain name configured, it would have to use
   full domain name. If optional domain list if defined, it will be
   sent to server as a hint. Defined in \cite{rfc3596}.

 \item[option ntp-server] -- (scope: interface). Takes optional address
   list as parameter. This option defines information about available NTP
   servers. Network Time Protocol \cite{rfc2030} is a protocol used
   for time synchronisation, so all hosts in the network has the same
   proper time set. If address list is specified, it will be sent to
   server as a hint. Defined in \cite{rfc4075}.

 \item[option time-zone] -- (scope: interface). Optionally takes one
   string parameter that specifies requested timezone. It is possible
   to retrieve timezone from the server. If client is interested in
   this information, it should ask for this option. Note that this
   option is considered obsolete as it is mentioned in draft version
   only \cite{draft-timezone}. Work on this draft seems to be
   abandoned as similar functionality is provided in now
   standard \cite{rfc4075}. Unfortunately, it is not supported in
   Dibbler yet.

 \item[option sip-server] -- (scope: interface). Takes optional
   address list as parameter. Session Initiation Protocol
   (SIP) \cite{rfc3263} is a control protocol for creating, modifying,
   and terminating sessions with one or more participants. These
   sessions include Internet telephone calls, multimedia distribution,
   and multimedia conferences, including VoIP phones. If address list
   is specified, it will be sent as a hint for a server. Format of
   this option is defined in \cite{rfc3319}.

 \item[option sip-domain] -- (scope: interface). Takes optional list
   of domains. It is possible to define domain names for Session
   Initiation Protocol \cite{rfc3263}. Configuration of this parameter
   will ease usage of domain names in the SIP protocol. If domain list
   is specified, it will be sent as a hint for a server. Format of
   this option is defined in \cite{rfc3319}.

 \item[option nis-server] -- (scope: interface). Takes optional
   address list as parameter. Network Information Service (NIS) is a
   Unix-based system designed to use common login and user information
   on multiple systems, e.g. universities, where students can log on
   to ther accounts from any host. To use this functionality, a host
   needs information about NIS server's address. This can be retrieved
   with this option. If address list is specified, it will be sent as
   a hint for a server. Its format is defined in \cite{rfc3898}.

 \item[option nis-domain] -- (scope: interface). Takes optional list
   of domains. Network Information Service (NIS) can albo specify
   domain names. It can be configured with this option. If domain list
   is specified, it will be sent as a hint for a server. It is defined
   in \cite{rfc3898}.

 \item[option nis+-server] -- (scope: interface). Takes optional
   address list as parameter. Network Information Service Plus (NIS+)
   is an improved version of the NIS protocol. If address list is
   specified, it will be sent as a hint for a server. This option is
   defined in \cite{rfc3898}.

 \item[option nis+-domain] -- (scope: interface). Takes optional list
   of domains. Similar to nis-domain, it defines domains for NIS+. If
   domain list is specified, it will be sent as a hint for a
   server. This option is defined in \cite{rfc3898}.

 \item[option lifetime] -- (scope: interface). This statement can be
   present or missing. The default is missing.
   Base spec of the DHCPv6 protocol does offers way of refreshing
   addresses only, but not the options. Lifetime defines, how often
   client would like to renew all its options. By default client will
   not send such option, but it will accept it and act accordingly if
   the server sends it on its own. Format of this option is defined
   in \cite{rfc4242}.

 \item[option aftr] -- (scope: interface). In networks that deploy
   Dual-Stack Lite architecture \cite{rfc6333}, client (B4) needs to
   configure DS-Lite tunnel. Client may obtain information about AFTR
   (a remote tunnel endpoint). This option conveys fully qualified
   domain name. This statement instructs client to request such
   option. It is defined in defined in \cite{rfc6334}.

\item[option] -- (scope: interface). There are number of options
   supported by Dibbler. There may be cases, however, when user wants
   to specify its own options. Several syntaxes are supported:
\begin{lstlisting}
option number - hexstring
option number address-keyword address
option number address-list
option number string-keyword string
option number address-keyword request-keyword
option number string request-keyword
option number address-list request-keyword
\end{lstlisting}
where number designates option number, address-keyword is word ``address'',
address is an IPv6 address, address-list is coma separated list of
IPv6 addresses, string-keyword is a word ``string'' and
request-keyword is a word ``request''. See Section \ref{feature-custom-options}.

 \item[auth-enabled] -- (scope: global). Takes one boolean
   parameter. The default is 0. This option enables authentication and
   authorization. When set to true, server support must also be
   enabled, otherwise all messages will be dropped. See
   Section \ref{feature-auth} for details.

 \item[auth-accept-methods] -- (scope: global). Takes coma separated
   list of accepted authentication methods methods that client will
   accept from server. If this list is empty, any method will be
   accepted. Possible values
   are: \texttt{none}, \texttt{digest\_plain}, \texttt{digest\_hmac\_md5},
   \texttt{digest\_hmac\_sha1}, \texttt{digest\_hmac\_sha224},
   \texttt{digest\_hmac\_sha256}, \texttt{digest\_hmac\_sha384},
   and \texttt{digest\_hmac\_sha512}.

\end{description}

After receiving options values from a server, client stores values of
those options in separate files in the working directory
(\verb+/var/lib/dibbler+ in POSIX systems (Linux, Mac OS X and BSD)
and current directory in Windows). File names start with the option
word, e.g. \verb+option-dns-server+. Dibbler client can also call user
defined script after parameters are assigned or removed. Dibbler
client also sets DNS servers and domain names on its own on most
systems.

\subsubsection{Stateless configuration}

If interface does not contain \verb+IA+ or \verb+TA+ keywords, client
will ask for one address (one IA with one address request will be sent).
If client should not request any addresses on this interface,
\opt{stateless}\footnote{In the version 0.2.1-RC1 and earlier, this
  directive was called no-ia. This depreciated name is valid for now,
  but might be removed in future releases.} keyword must be used. In
such circumstances, only specified options will be requested.

\subsubsection{Relay support}
Usage of the relays is not visible from the client's point of view:
Client can't detect if it communicates via relay(s) or directly
with the server. Therefore no special directives on the client side
are required to use relays. See section \ref{feature-relays} for
details related to relay deployment.

\subsection{Client configuration examples}
This subsection contains various examples of the most popular
configurations. Several additional examples are provided with the source
code. Please download it and look at \verb+*.conf+ files.

\subsubsection{Example 1: Default}
In the most simple case, client configuration file can be empty. Client will try to
assign one address for every interface present in the system, except
interfaces, which are:
\begin{itemize}
\item down (flag UP not set)
\item loopback (flag LOOPBACK set)
\item not running (flag RUNNING not set)
\item not multicast capable (flag MULTICAST not set)
\item have link-layer address less than 6 bytes long (this requirement
      should skip all tunnels and virtual interfaces)
\end{itemize}

If you must use DHCPv6 on one of such interfaces (which is not
recommended and such attempt probably will fail), you must explicitly
specify this interface in the configuration file.

\subsubsection{Example 2: DNS}
Configuration mentioned in previous subsection is a minimal one and in a
real life will be used rarely. The most common usage of the DHCPv6
protocol is to request for an address and DNS configuration. Client
configuration file achieving those goals is presented below:
\begin{lstlisting}
# client.conf
log-mode short
log-level 7
iface eth0 {
  ia
  option dns-server
}
\end{lstlisting}

\subsubsection{Example 3: Timeouts and specific address}

Automatic configuration is being driven by several timers, which define,
what action should be performed at various intervals. Since all
values are provided by the server, client can only define values, which
will be sent to a server as hints. Server might take them into
consideration, but might also ignore them
completely. Following example shows how to ask for a specific address
and provide hints for a server. Client would like to get 2000::1:2:3
address, it would like to renew addresses once in 30 minutes (T1 timer
is set to 1800 seconds). Client also would like to have address, which
is prefered for an hour and is valid for 2 hours.

\begin{lstlisting}
# client.conf
log-mode short
log-level 7
iface eth0 {
  T1 1800
  T2 2000
  prefered-lifetime 3600
  valid-lifetime 7200
  ia {
    address {
      2000::1:2:3
    }
  }
}
\end{lstlisting}

\subsubsection{Example 4: More than one address}

Another example: client would like to obtain 2 addresses on
wifi0 interface. They are necessary since this particular interface name
contains spaces. It is possible to do this in two ways. First is to
sent 2 Identity Associations (IA for short). Identity Association is a
nice name for a addresses container. This appears to be a most common
way of telling server that this client is interested in more than one
address.

\begin{lstlisting}
# client.conf
log-mode short
log-level 5
iface wifi0 {
  ia
  ia
}
\end{lstlisting}

Another way it to send one IA, but include two address hints in
it. Server may take them into consideration (dibbler server does), but
some other DHCPv6 implementations may ignore those hints.

\begin{lstlisting}
# client.conf
log-mode short
log-level 5
iface wifi0 {
  ia {
     address
     address
  }
}
\end{lstlisting}

\subsubsection{Example 5: Quick configuration using Rapid-commit}
Rapid-commit is a shortened exchange with server. It consists of only
two messages, instead of the usual four. It is worth to know that both sides (client
and server) must also support rapid-commit to use this fast
configuration.

\begin{lstlisting}
# client.conf
iface eth1 {
  rapid-commit yes
  ia
  option dns-server
}
\end{lstlisting}

\subsubsection{Example 6: Stateless mode}
Client can be configured to work in a stateless mode. It means that it
will obtain only some configuration parameters, but no
addresses. Let's assume we want all the details stored in a log file and
we want to obtain all possible configuration parameters. Here is a
configuration file:

\begin{lstlisting}
# client.conf
log-level 8
log-mode full
stateless
iface eth0
{
  option dns-server
  option domain
  option ntp-server
  option time-zone
  option sip-server
  option sip-domain
  option nis-server
  option nis-domain
  option nis+-server
  option nis+-domain
}
\end{lstlisting}

\subsubsection{Example 7: Dynamic DNS (FQDN)}
\label{example-client-fqdn}
Dibbler client is able to request fully qualified domain name,
i.e. name, which is fully resolvable using DNS. After receiving such
name, it can perform DNS Update procedure. Client can ask for any
name, without any preferrence. Here is an example how to configure
client to perform such task:
\begin{lstlisting}
# client.conf

# Set protocol to one of the following values: udp, tcp, any
ddns-protocol udp

# Sets DDNS Update timeout (in ms)
ddns-timeout 800

# uncomment following line to force S bit to 0
# option fqdn-s 0
log-level 7

iface eth0 {
# ask for one address
    ia

# ask for options
   option dns-server
   option domain
   option fqdn

# ask for fully qualified domain name (any name will do)
  option fqdn
# you can also provide hint for the server regarding preferred name
#  option fqdn dexter.example.org

}
\end{lstlisting}

In this case, client will mention that it is interested in FQDN by
using Option Request and empty FQDN option, as specified in
\cite{rfc4704}. Server upon receiving such request (if it is
configured to support it), will provide FQDN option containing domain
name. Depending on the server's configuration, all DNS Updates will be
performed by the server, forward will be performed by client and reverse
by the server, or only forward will be done by a client.

It is also possible for client to provide its name as a hint for
server. Server might take it into consideration when it will choose a
name for this client. To send specific hostname, additional parameter
(a string with a fully qualified domain name) should be specified.

Two additional parameters were introduced in Dibbler
0.8.1. \verb+ddns-protocol+ specifies protocol that should be used for
communication with DNS server.  Allowed values
are \verb+udp+, \verb+tcp+ or \verb+any+. ``Any'' will try to use UDP
and if that fails, it will revert to TCP. Second parameter
is \verb+ddns-timeout+ that specifies maximum time allowed for DNS
server to respond before assuming communication failure. It is
specified in milliseconds.

Note that to successfully perform DNS Update, address must be assigned
and dns server address must be known. Therefoe  ``ia'' and ``option
dns-server'' are required for ``option fqdn'' to work properly. Also if
DHCPv6 server provides more than one DNS server address, update will
be attempted only for the first address on the list.

It is also possible to force S bit in the FQDN option to 0 or 1. See
\cite{rfc4704} for details regarding its meaning.

\subsubsection{Example 8: Interface indexes}
Usually, interface names are referred to by names, e.g. eth0 or Local
Area Connection. Every system also provides unique number associated
with each infterface, usually called ifindex or interface index. It is
possible to read the number using \verb+ip l+ command (Linux) or
\verb+ipv6 ifx+. Below is an example, which demonstrate how to use
interface indexes:

\begin{lstlisting}
# client.conf
log-mode short
log-level 5
iface 5 {
  ia
}
\end{lstlisting}

\subsubsection{Example 9: Vendor-specific options}
\label{example-client-vendor-spec}
It is possible to configure dibbler-client to ask for a vendor specific
options. Although there are no dibbler-specific features to configure,
it is possible to use this option to test other server
implementations. This option will rather be used by network engineers
and power network admins, rather than normal end users.

There are 3 ways to define, how dibbler-client can request
vendor-specific options. First choice: It can just ask for this option (only
\opt{option request option} will be sent). Second choice: it can ask for
vendor-spec option by adding such option with enterprise number set, but
no actual data. Third choice: send this option and include both
enterprise number and actual data. In the following configuration file
example, uncomment appropriate line to obtain desired bahavior:

\begin{lstlisting}
# client.conf
log-level 8
iface eth0 {
# ask for address
    ia

# uncomment only one of the following lines:
   option vendor-spec
#   option vendor-spec 1234
#   option vendor-spec 1234 0x0002abcd

# To ask for multiple vendor-spec options, uncomment:
#   option vendor-spec 123,456
}
\end{lstlisting}

Although that is almost never needed, it is possible to configure
client to request multiple vendor-specific options at the same
time. That feature is mainly used as a test tool for the server. To
use it, uncomment last line in the example above.

\subsubsection{Example 10: Unicast communication}

Client would like to obtain an address on ,,Local Area Connection''
interface. Note quotation marks around
interface name. They are necessary since this particular interface name
contains spaces. Client also would like to accept Unicast
communication if server supports it. User wants all information
to be logged via Linux syslog daemon. Take note that you won't be
able see to what Dibbler is doing with such low log-level. (Usually
log-level should be set to 7, which is also a default value).

\begin{lstlisting}
# client.conf
log-mode syslog
log-level 5
iface "Local Area Connection" {
  unicast yes
  ia
  ia
}
\end{lstlisting}

\subsubsection{Example 11: Prefix delegation}
\label{example-client-prefix}
From the client's point of view, configuration is quite simple. It is
required to specify that this client is interested in prefix
delegation. See section \ref{feature-prefix} for background
information related to prefix delegation and sections
\ref{example-server-prefix} and \ref{example-server-prefixes} for
details about server configuration. To ask for prefix delegation,
emph{prefix-delegation} (or \emph{pd}) should be used.

\begin{lstlisting}
# client.conf
iface "eth0" {
  ia  // ask for address
  pd  // ask for prefix
}
\end{lstlisting}

It is possible to define additional parameters for a prefix:

\begin{lstlisting}
# client.conf
iface eth0 {
  pd {
    t1 1000
    t2 2000
  }
}
\end{lstlisting}

\subsubsection{Example 12: Insist mode}
\label{example-client-insistmode}
During normal operation, when client asks for an option, but does not
receive it from the server, it complain, but takes no action. To force
client to insist (i.e. ask over and over again), so called insist mode
has been introduced. See section \ref{feature-insist-mode} for
extended explanation.

\begin{lstlisting}
insist-mode
iface "eth0" {
   ia
   option dns-server
   option domain
   option ntp-server
}
\end{lstlisting}

\subsubsection{Example 13: Inactive mode}
\label{example-client-inactivemode}
Usually client starts when network interfaces are
operational. Normally downed or nonexisting interfaces mentioned in
the configuration file are considered misconfiguration and client
refuses to start. However, sometimes that is not the case, e.g. still
waiting to be associated wireless interfaces. To allow operation in
such circumstances, inactive mode has been added. See
\ref{feature-inactive-mode} for detailed explanation.
interfaces are spec

\begin{lstlisting}
inactive-mode
iface "eth0" {
  ia
}
\end{lstlisting}

\subsubsection{Example 14: Authentication}
\label{example-client-auth}
Authentication is enabled. Client will accept HMAC-SHA-512, HMAC-MD5
and HMAC-SHA-256 as an authentication method.

\begin{lstlisting}
# client.conf

log-mode short
log-level 7

auth-enabled true
auth-accept-methods digest-hmac-sha512, digest-hmac-md5, digest-hmac-sha256

iface eth0 {
}
\end{lstlisting}

\subsubsection{Example 15: Skip Confirm}
\label{example-client-confirm}
Client detects if previous client instance was not shutdown properly
(due to power outage, client crash or similar event). In such case, it
reads existing address database and checks if assigned addresses may
still be valid. If that is so, it tries to confirm those addresses by
using \msg{CONFIRM} message.

If user don't want \msg{CONFIRM} message to be send and client should
start ''from scratch'' every time, it is possible to disable confirm
support.

\begin{lstlisting}
# client.conf

log-mode short
log-level 7
skip-confirm

iface eth0 {
  ia
}
\end{lstlisting}

\subsubsection{Example 15: User-defined IAID}
\label{example-client-iaid}
Sometimes it is useful to define specific IAID  identifiers. That is
rather uncommon, but possible. This technique can be used for both
addresses (IA\_NA options) and prefixes (IA\_PD options).

\begin{lstlisting}
# client.conf

iface "eth0" {
  ia 123
  option dns-server
  option domain
}
\end{lstlisting}

\subsubsection{Example 16: DS-Lite tunnel (AFTR)}
\label{example-server-ds-lite}
Server may provide information about AFTR (a Dual Stack Lite tunnel
endpoint) to the clients, as specified in \cite{rfc6334}.

\begin{lstlisting}
iface "eth0" {
  ia
  option aftr # request name of the remote DS-Lite tunnel endpoint
}
\end{lstlisting}

\subsubsection{Example 17: Custom options}
Client is able to ask for custom options, that are not supported by
default. Following config file allows client to ask for many
options. Also, see Section \ref{feature-custom-options} for
extended explanation.

\begin{lstlisting}
#client.conf
iface "eth0" {
  unicast 1
  ia

  option 145 - 01:02:a3:b4:c5:dd:ea
  option 146 address 2001:db8:1::dead:beef
  option 147 address-list 2001:db8:1::aaaa,2001:db8:1::bbbb
  option 148 string "secretlair.example.org"

  option 149 string request
  option 150 address request
  option 151 address-list request
\end{lstlisting}

\subsubsection{Example 18: Remote Autoconfiguration}
Client is able to use experimental extension to ask for configuration
remotely. See Section \ref{feature-remote-autoconf} for details.

\begin{lstlisting}
log-mode short
log-level 8
experimental
remote-autoconf

iface "eth0" {
  ia
  unicast 1
  option dns-server
  option domain
  option nis-server
  option nis-domain
  option nis+-server
  option nis+-domain
  option time-zone
  option lifetime
}
\end{lstlisting}

\subsection{Server configuration file}
\label{server-conf}
Server configuration is stored in \verb+server.conf+ file in the
\verb+/etc/dibbler+ (Linux systems) or in current (Windows systems)
directory.

\subsubsection{Global scope}

Every option can be declared in a global scope. Global options can be
defined here. Also options of a smaller scopes can be defined here --
they will be used as a default values. Configuration file has following syntax:

\begin{lstlisting}
 global-options
 interface-options
 class-options
 interface-declaration
\end{lstlisting}

\subsubsection{Interface declaration}
Each network interface, which should be serviced by the server, must be
mentioned in the configuration file. Network interface is defined like this:
\begin{lstlisting}
iface interface-name
{
  interface-options
  class-options
}
\end{lstlisting}

or

\begin{lstlisting}
iface number
{
  interface-options
  class-options
}
\end{lstlisting}

where \verb+interface-name+ denotes name of the interface and
\verb+interface-number+ denotes its number. Name no longer needs to be
enclosed in single or double quotes (except Windows systems, when
interface name contains spaces). Note that virtual interfaces, used
to setup relay support are also declared in this way.

\subsubsection{Class scope}
Class is a smallest scope used in the server configuration file. It
contains definition of the addresses, which will be provided to
clients. Only class scoped parameters can be defined here. Address class
is declared as follows:
\begin{lstlisting}
class
{
  class-options
  address-pool
}
\end{lstlisting}

Address pool defines range of the addresses, which can be assigned to the
clients. It can be defined in one of the following formats:
\begin{lstlisting}
pool minaddress-maxaddress
pool address/prefix
\end{lstlisting}

\subsubsection{Standard options}

So called standard options are defined by the base DHCPv6 specification,
a so called RFC 3315 document \cite{rfc3315}. Those options are
called standard, because all DHCPv6 implementations, should properly
handle them. Each option has a specific scope it belongs to.

Standard options are declared in the following way:

\begin{lstlisting}
OptionName option-value
\end{lstlisting}

\begin{description}
 \item[work-dir] -- (scope: global, type: string, default: .) Defines working
            directory.
 \item[log-level] -- (scope: global, type: integer, default: 7) Defines
            verbose level of the log messages. The valid range if
            from 1 (Emergency) to 8 (Debug). The higher the logging
            level is set, the more messages dibbler will print.
 \item[log-name] -- (scope: global, type: string, default: Server). Defines
            name, which should be used during logging.
 \item[log-mode] -- (scope: global, type: short, full, precise or syslog
            default value: full) Defines logging mode. In the
            default, full mode, name, date and time in the h:m:s format
            will be printed. In short mode, only minutes and
            seconds will be printed (this mode is useful on
            terminals with limited width). Recently added precise
            mode logs information with seconds and microsecond
            precision. It is a useful for finding bottlenecks in
            the DHCPv6 autoconfiguration process. Syslog is a Linux
            mode only.
 \item[cache-size] -- (scope: global, type: integer, default:
            1048576). It defines a size of the memory (specified in
            bytes) which can se used to store cached entries.
\item[interface-id-order] -- (scope: global, type: before, after or omit,
        default: before) This parameter defines placement of the
        interface-id option. During message relaying options can be placed in the \msg{RELAY-REPL}
        message is arbitrary order. This option has been specified to control
        that order. \opt{interface-id} option can be placed before or after
        \opt{relay-message} option. There is also possibility to instruct
        server to omit the \opt{interface-id} option altogether, but since
        this violates \cite{rfc3315}, it should not be used. In general, this
        configuration parameter is only useful when dealing with buggy relays,
        which can't handle all option orders properly. Consider this parameter
        a debugging feature. Note: similar parameter is available in the dibbler-relay.
\item[inactive-mode] -- (scope: global, type: present or missing,
  default: missing). This enables so called inactive mode. When server
  begins operation and it detects that required interfaces are not
  ready, error message is printed and server exits. However, if
  inactive mode is enabled, server sleeps instead and wait for
  required interfaces to become operational. That is a useful feature,
  when using wireless interfaces, which take some time to initialize
  as associate.
\item[guess-mode] -- (scope: global, type: present or missing,
  default: missing). When this option is enabled, server will not pay
  close attention to the interface-id option in relayed messages. If
  interface-id has a value other than specified in server.conf or even
  when there is no interface-id option at all, it will use first relay
  defined.
 \item[preference] -- (scope: interface, type: 0-255, default:
            none). Eech server can be configured to a specific
            preference level. When client receives several
            \msg{ADVERTISE} messages, it should choose that server,
            which has the highest preference level. It is also worth
            noting that client, upon reception of the \msg{ADVERTISE}
            message with preference set to 255 should skip wait phase
            for possible other \msg{ADVERTISE} messages.
 \item[unicast] -- (scope: interface, type: address,
            default:none). Normally clients sends data to a well known
            multicast address. This is easy to achieve, but it wastes
            network resources as all nodes in the network must process
            such messages and also network load is increased. To prevent
            this, server might be configured to inform clients about its
            unicast address, so clients, which accept it, will switch to
            a unicast communication.
 \item[rapid-commit] -- (scope: interface, type: boolean, default:
            0). This option allows rapid commit procedure to be
            performed. Note that enabling rapid commit on the server
            side is not enough. Client must be configured to allow
            rapid commit, too.
\item[iface-max-lease] -- (scope: interface, type: integer, default:
            $2^{32}-1$). This parameter defines, how many normal
            addresses can be granted on this interface.
\item[client-max-lease] -- (scope: interface, type: interger,
            default:$2^{32}-1$). This parameter defines, how many
            addresses one client can get. Main purpose of this
            parameter is to limit number of used addresses by
            misbehaving (malicious or restarting) clients.
\item[relay] -- (scope: interface, type: string, default: not
            defined). Used in relay definition.
            It specifies name of the physical (or name of
            another relay, if cascade relaying is used) interface,
            which is used to receive and transmit relayed data. See
            \ref{feature-relays} for details of relay deployment and
            sections \ref{example-server-relay1} and
            \ref{example-server-relay2} for configuration examples.
\item[interface-id] -- (scope: interface, type: integer, default: not
            defined). Used in relay definition. Each relay interface
            should have defined its unique identified. It will be sent
            in the \opt{interface-id} option. Note that this value
            must be the same as configured in the dibbler-relay. It
            may be possible to specify this parameter by using a
            number (option will be 4 bytes long), a string or a full
            hex dump.  See
            \ref{feature-relays}, \ref{example-server-relay1} and
            \ref{example-relay} for details.
 \item[vendor-spec] -- (scope: interface, type: integer-hexstring,
            default: not defined). This parameter can be used to
            configure some vendor-specific information option. Since
            there are no dibbler-specific options, this implementation
            is flexible. User can specify in the configuration file,
            how should this option look like. See
            \ref{example-server-vendor-spec} section for details. It
            is uncommon, but possible to define several vendor
            specific options for different vendors. In such case,
            administrator must specify coma separated list. Each list
            entry is a vendor (enterprise number), ,,--'' sign and a
            hex dump (similar to DUID).
 \item[T1] -- (scope: class, type: integer or integer range: default:
            $2^{32}-1$). This value
            defines after what time client should start renew
            process. Exact value or accepted range can be
            specified. When exact value is defined, client's hints are
            ignored completely.
 \item[T2] -- (scope: class, type: integer or integer range, default:$2^{32}-1$). This value
            defines after what time client will start emergency rebind
            procedure if renew process fails. Exact value or accepted range can be
            specified. When exact value is defined, client's hints are
            ignored completely.
\item[valid-lifetime] (scope: class, type: integer or integer range,
            default:$2^{32}-1$). This parameter defines valid lifetime of
            the granted addresses. If range is specified, client's
            hints from that range are accepted.
\item[prefered-lifetime] (scope: class, type: integer or integer range,
            default:$2^{32}-1$). This parameter defines prefered
            lifetime of the granted addresses. If range is specified,
            client's hits from that range will be accepted.
\item[class-max-lease]  -- (scope: interface, type: interger,
            default:$2^{32}-1$). This parameter defines, how many
            addresses can be assigned from that class.
\item[reject-clients] -- (scope: class, type: address or DUID list,
            default: none). This parameter is sometimes called
            black-list. It is a list of a clients, which should not be
            supported. Clients can be identified by theirs link-local
            addresses or DUIDs.
\item[accept-only] -- (scope: class, type: address or DUID list,
            default: none). This parameter is sometimes called
            white-list. It is a list of supported clients. When this
            list is not defined, by default all clients (except
            mentioned in reject-clients) are supported. When
            accept-only list is defined, only client from that list
            will be supported.
\end{description}

\subsubsection{Additional options}
Server supports additional options, not specified in \cite{rfc3315}. They have
following generic form:

\begin{lstlisting}
option OptionName OptionsValue
\end{lstlisting}

All supported options are specified below:

\begin{description}
 \item[dns-server] -- (scope: interface, type: address list, default:
   none). This option conveys information about DNS servers
   available. After retriving this information, clients will be able
   to resolve domain names into IP (both IPv4 and IPv6)
   addresses. Defined in \cite{rfc3596}.
 \item[domain] -- (scope: interface, type: domain list, default:
   none). This option is used for configuring one or more domain
   names, which clients are connected in. For example, if client's
   hostname is \verb+alice.mylab.example.com+ and it wants to contact
   \verb+bob.mylab.example.com+, it can simply refer to it as
   \verb+bob+. Without domain name configured, it would have to use
   full domain name. Defined in \cite{rfc3596}.
 \item[ntp-server] -- (scope: interface, type: address list, default:
   none). This option defines information about available NTP
   servers. Network Time Protocol \cite{rfc2030} is a protocol used
   for time synchronisation, so all hosts in the network has the same
   proper time set. Defined in \cite{rfc4075}.
 \item[time-zone] -- (scope: interface, type: timezone, default:
   none). It is possible to configure timezone, which is provided by
   the server. Note that this option is considered obsolete as it is
   mentioned in draft version only \cite{draft-timezone}. Work on this
   draft seems to be abandoned as similar functionality is provided by
   now standard \cite{rfc4075}.
 \item[sip-server] -- (scope: interface, type: address list, default:
   none). Session Initiation Protocol \cite{rfc3263} is an control
   protocol for creating, modifying, and terminating sessions with one
   or more participants. These sessions include Internet telephone
   calls, multimedia distribution, and multimedia conferences. Its
   most common usage is VoIP. Format of this option is defined in
   \cite{rfc3319}.
 \item[sip-domain] -- (scope: interface, type: domain list, default:
   none). It is possible to define domain names for Session Initiation
   Protocol \cite{rfc3263}. Configuration of this parameter will ease
   usage of domain names in the SIP protocol. Format of this option is
   defined in \cite{rfc3319}.
 \item[nis-server] -- (scope: interface, type: address list, default:
   none). Network Information Service (NIS) is a Unix-based system
   designed to use common login and user information on multiple
   systems, e.g. universities, where students can log on to ther
   accounts from any host. Its format is defined in \cite{rfc3898}.
 \item[nis-domain] -- (scope: interface, type: domain list, default:
   none). Network Information Service (NIS) can albo specify domain
   names. It can be configured with this option. It is defined in
   \cite{rfc3898}.
 \item[nis+-server] -- (scope: interface, type: address list, default:
   none). Network Information Service Plus (NIS+) is an improved
   version of the NIS protocol. This option is defined in
   \cite{rfc3898}.
 \item[nis+-domain] -- (scope: interface, type: domain list, default:
   none). Similar to nis-domain, it defines domains for NIS+. This
   option is defined in \cite{rfc3898}.
 \item[lifetime] -- (scope: interface, type: boolean, default:
   no). Base spec of the DHCPv6 protocol does offers way of refreshing
   addresses only, but not the options. Lifetime defines, how often
   client should renew all its options. When defined, lifetime option
   will be appended to all replies, which server sends to a client. If
   client does not support it, it should ignore this option. Format of
   this option is defined in \cite{rfc4242}.
 \item[aftr] -- (scope: interface, type: FQDN). In Dual-Stack Lite
   networks, client may want to configure DS-Lite tunnel. Client may
   want to obtain information about AFTR (a remote tunnel
   endpoint). This option conveys a fully qualified domain name of the
   remote tunnel. This option is defined in \cite{rfc6334}.
 \item[auth-method] -- (scope: global, type: string, default:
   empty). Set it to one of the following values to enable
   authentication on ther server side, using selected method of
   generating authentication information:
   \begin{itemize}
   \item \texttt{none}
   \item \texttt{digest\_plain}
   \item \texttt{digest\_hmac\_md5}
   \item \texttt{digest\_hmac\_sha1}
   \item \texttt{digest\_hmac\_sha224}
   \item \texttt{digest\_hmac\_sha256}
   \item \texttt{digest\_hmac\_sha384}
   \item \texttt{digest\_hmac\_sha512}
   \end{itemize}
 \item[auth-lifetime] -- (scope: global, type: integer, default:
   0). Authentication lifetime. Currently not supported.
 \item[auth-key-len] -- (scope: global, type: integes, default:
   32). Key generation nonce length (see \cite{draft-aaa} for
   details).
\end{description}

Lifetime is a special case. It is not set up by client in a system
configuration. It is, however, used by the client to know how long
obtained values are correct and initiate \msg{RENEW} or
\msg{INF-REQUEST} message exchange to refresh received options.

\subsection{Server configuration examples}

This subsection contains various examples of the server
configuration. If you are interested in additional examples, download
source version and look at \verb+*.conf+ files.

\subsubsection{Example 1: Simple}

In opposite to client, server uses only interfaces described in config
file. Let's examine this common situation: server has interface named
\emph{eth0} (which is fourth interface in the system) and is supposed
to assign addresses from 2000::100/124 class. Simplest config file
looks like that:

\begin{lstlisting}
# server.conf
iface eth0
{
  class
  {
    pool 2000::100-2000::10f
  }
}
\end{lstlisting}

\subsubsection{Example 2: Timeouts}
Server should be configured to deliver specific timer values to the
clients. This example shows how to instruct client to renew (T1 timer)
addresses one in 10 minutes. In case of problems, ask other servers in
15 minutes (T2 timer), that allowe prefered lifetime range is from 30
minutes to 2 hours, and valid lifetime is from 1 hour to 1 day. DNS
server parameter is also provided. Lifetime option is used to make
clients renew all non-address related options renew once in 2 hours.

\begin{lstlisting}
# server.conf
iface eth0
{
  T1 600
  T2 900
  prefered-lifetime 1800-3600
  valid-lifetime 3600-86400
  class
  {
    pool 2000::100/80
  }

  option dns-server 2000::1234
  option lifetime 7200
}
\end{lstlisting}

\subsubsection{Example 3: Limiting amount of addresses}
Another example: Server should support 2000::0/120 class on eth0
interface. It should not allow any client to obtain more than 5
addresses and should not grant more then 50 addresses in total. From
this specific class only 20 addresses can be assigned. Server
preference should be set to 7. This means that this server is more
important than all server with preference set to 6 or less.
Config file is presented below:

\begin{lstlisting}
# server.conf
iface eth0
{
  iface-max-lease 50
  client-max-lease 5
  preference 7
  class
  {
    class-max-lease 20
    pool 2000::1-2000::100
  }
}
\end{lstlisting}

\subsubsection{Example 4: Unicast communication}
\label{example-server-unicast}

Here's modified previous example. Instead of specified limits, unicast
communication should be supported and server should listen on
2000::1234 address. Note that default multicast address is still
supported. You must have this unicast address already configured on
server's interface.

\begin{lstlisting}
# server.conf
log-level 7
iface eth0
{
  unicast 2000::1234
  class
  {
    pool 2000::1-2000::100
  }
}
\end{lstlisting}

\subsubsection{Example 5: Rapid-commit}
This configuration can be called quick. Rapid-commit is a way to shorten exchange to only two messages. It is
quite useful in networks with heavy load. In case if client does not
support rapid-commit, another trick is used. Preference is set to
maximum possible value. 255 has a special meaning: it makes client to
skip wait phase for possible advertise messages from other servers and
quickly request addresses.

\begin{lstlisting}
# server.conf
log-level 7
iface eth0
{
  rapid-commit yes
  preference 255
  class
  {
    pool 2000::1/112
  }
}
\end{lstlisting}

\subsubsection{Example 6: Access control}
Administrators can selectively allow certain client to use this
server (white-list). On the other hand, some clients could be
explicitly forbidden to use this server (black-list). Specific DUIDs,
DUID ranges, link-local addresses or the whole address ranges are
supported. Here is config file:

\begin{lstlisting}
# server.conf
iface eth0
{
  class
  {
    # duid of the rejected client
    reject-clients ``00001231200adeaaa''
    2000::2f-2000::20  // it's in reverse order, but it works.
                       // just a trick.
  }
}
iface eth1
{
  class
  {
    accept-only fe80::200:39ff:fe4b:1abc
    pool 2000::fe00-2000::feff
  }
}
\end{lstlisting}

\subsubsection{Example 7: Multiple classes}
Although this is not common, a few users have requested support for multiple classes on one interface.
Dibbler server can be configured to use several classes. When client asks for an address, one of the classes
is being choosen on a random basis. If not specified otherwise, all classes have equal probability of being chosen.
However, this behavior can be modified using \verb+share+ parameter. In the following example, server supports
3 classes with different preference level: class 1 has 100, class 2 has 200 and class 3 has 300. This means that class 1
gets $\frac{100}{100+200+300} \approx 16\% $ of all requests, class 2
gets $\frac{200}{100+200+300} \approx 33\% $ and class 3 gets the rest
($\frac{300}{100+200+300}=50\% $).

\begin{lstlisting}
# server.conf
log-level 7
log-mode short

iface eth0 {
 T1 1000
 T2 2000

 class {
   share 100
   pool 4000::1/80
 }
 class {
   share 200
   pool 2000::1-2000::ff
 }

 class {
   share 300
   pool 3000::1234:5678/112
 }
}
\end{lstlisting}

\subsubsection{Example 8: Relay support}
\label{example-server-relay1}
To get more informations about relay configuration, see section \ref{feature-relays}.
Following server configuration example explains how to use
relays. There is some remote relay with will send encapsulated data over
eth1 interface. It is configured to append interface-id option set to
5020 value. Let's allow all clients using this relay some addresses
and information about DNS servers. Also see section
\ref{example-relay-1} for corresponding relay configuration.

Note that although eth1 interface is mentioned in the configuration file,
direct traffic from clients located on the eth1 interface will not be
supported. In this example, eth1 is used only to support requests
relayed from remote link identified with interface-id value 5020.
Of course it is possible to support both local and remote traffic. In
such case, normal eth1 definition should be present in the server
configuration file. Also note that real (physical) interfaces should
be specified before logical ones.

\begin{lstlisting}
# server.conf
iface relay1 {
  relay eth1
  // interface-id 5020
  // interface-id "some interface name"
  interface-id 0x427531264361332f3000001018680f980000

  class {
    pool 2000::1-2000::ff
  }
  option dns-server 2000::100,2000::101
}
\end{lstlisting}

\subsubsection{Example 9: Cascade 2 relays}
\label{example-server-relay2}
This is an advanced configuration. It assumes that client sends data to
relay1, which encapsulates it and forwards it to relay2, which
eventually sends it to the server (after additional encapsulation). It
assumes that first relay adds interface-id option set to 6011 and
second one adds similar option set to 6021. For details about relays
in general and cascade setup in particular, see section
\ref{feature-relays}. Also see section \ref{example-relay-cascade}
for corresponding relays configuration.

\begin{lstlisting}
# server.conf
iface relay1
{
  relay eth0
  interface-id 6011
}

iface relay2
{
  relay relay1
  interface-id 6021
  T1 1000
  T2 2000
  class {
    pool 6020::20-6020::ff
  }
}
\end{lstlisting}

\subsubsection{Example 10: Dynamic DNS (FQDN)}
\label{example-server-fqdn}

Support for Dynamic DNS Updates was added in version 0.5.0RC1. To
configure it on the server side, list of available names usually
should be defined. Each name can be reserved for a certain address or
DUID. When no reservation is specified, it will available to everyone,
i.e. the first client asks for FQDN will get this name. In following
example, name 'zebuline.example.com' is reserved for address 2001::db8:1::1,
kael.example.com is reserved for 2001:db8:1::2 and test.example.com is
reserved for client using DUID
00:01:00:00:43:ce:25:b4:00:13:d4:02:4b:f5.

Also note that is required to define, which side can perform updates.
This is done using single number after ,,option fqdn'' phrase. Server
can perform two kinds of DNS Updates: AAAA (forward resolving,
i.e. name to address) and PTR (reverse resolving, i.e. address to
name). To configure server to execute both updates, specify 2. This is
a default behavior. If this value will be skipped, server will attempt
to perform both updates. When 1 will be specified, server will update
PTR record only and will leave updating AAAA record to the
client. When this value is set to 0, server will not perform any
updates.

The last parameter (64 in the following example) is a prefix length of
the reverse domain supported by the DNS server, i.e. if this is set to
64, and 2000::/64 addresses are used, DNS server must support
0.0.0.0.0.0.0.0.0.0.0.0.0.0.2.ip6.arpa. zone.

There are several additional parameters that affect DNS Update
mechanism. \verb+ddns-protocol+ specifies protocol that should be used
for communication with DNS server.  Allowed values
are \verb+udp+, \verb+tcp+ or \verb+any+. ``Any'' will try to use UDP
and if that fails, it will revert to TCP. Second parameter
is \verb+ddns-timeout+ that specifies maximum time allowed for DNS
server to respond before assuming communication failure. It is
specified in milliseconds.

The next useful parameter is \verb+fqdn-ddns-address+ that specifies
address of DNS server that updates should be performed to. If it is
not specified, first DNS address from \verb+option dns-server+ will be
used.

The last important parameter is +\verb+accept-unknown-fqdn+. In a
simplest scenario, server is configured with a list of allowed
names. Connecting clients may get only those names. That is convenient
case, but it is often not feasible to deploy it in a real network.  In
real networks, clients usually send out their own names, rather than
wait for server to assign them one. In such cases, it is somewhat
expected that server will not have complete list of all possible names
that clients may send. Thus sooner or later server will likely receive
a fqdn name that is unknown to him. This parameter specifies server
behavior in such case.

There are 5 possible policied currently supported. Each one is
identified with an integer between 0 and 4. 0 means that uknown names
should be rejected. That policy is useful for strictly controlled
networks. 1 means that other available name from list of possible
names should be sent instead. This is a compromise between strict
control over names and liberal acceptance of clients' names. Policy 2
accept any name that client will send. Names will be sanity
checked. Note that mobile and nomadic clients may send names from
their home networks. That may be a problem if server attempts to
update AAAA records as its DNS server will probably only accept AAAA
updates for locally administered domains. As a solution to this
problem, policy 3 was implemented. It takes client's hostname (without
its domain name), appends local domain name and uses such constructed
fully qualified domain name. For example, if client
sends \verb+nomad.faraway.org+ while visiting \verb+example.org+, with
this policy in place, \verb+nomad.example.org+ will be assigned.

The last policy is useful for larger networks. Instead of accepting
clients' ideas about their hostnames, dedicated name is generated
based on assigned address. For example, client that received
\verb+2001:db8:1:0:c7a8:e81:c500:46ce+ address in domain
\verb+example.org+ will be assigned
a \verb+2001-db8-1-0-c7a8-e81-c500-46ce.example.org+ name.

Following config file is a good starting point for tweaking DNS
Update enabled server configuration.

\begin{lstlisting}
# server.conf

# Logging level range: 1(Emergency)-8(Debug)
#
log-level 8

# Don't log full date
log-mode short

# Set protocol to one of the following values: udp, tcp, any
ddns-protocol udp

# Sets DDNS Update timeout (in ms)
ddns-timeout 1000

# specify address of DNS server to be used for DDNS
fqdn-ddns-address 2001::1

iface "eth0" {

# assign addresses from this class
 class {
   pool 2001:db8:1::/64
 }

# provide DNS server location to the clients
# also server will use this address to perform DNS Update,
# so it must be valid and DNS server must accept DNS Updates.
 option dns-server 2000::1

# provide their domain name
 option domain example.com

# provide fully qualified domain names for clients
# note that first, second and third entry is reserved
# for a specific address or a DUID
 option fqdn 1 64
             zebuline.example.com - 2000::1,
             kael.example.com - 2000::2,
             wash.example.com - 0x0001000043ce25b40013d4024bf5,
             zoe.example.com,
             malcolm.example.com,
             kaylee.example.com,
             jayne.example.com,
             inara.example.com

# specify what to do with client's names that are not on the list
# 0 - reject
# 1 - send other name from allowed list
# 2 - accept any name client sends
# 3 - accept any name client sends, but append specified domain suffix
# 4 - ignore client's hint, generate name based on his address, append domain name

 accept-unknown-fqdn 4 example.org

}
\end{lstlisting}

\subsubsection{Example 11: Vendor-specific Information option}
\label{example-server-vendor-spec}
It is possible to configure dibbler-server to provide vendor-specific
information options. Since there are no dibbler-specific parameters,
this implementation is quite flexible. Enterprise number as well as
content of the option itself can be configured.

\begin{lstlisting}
# server.conf
log-level 8
log-mode precise
iface "eth1" {
 class {
   pool 2000::1-2000::ff
 }

 option vendor-spec 1234-0x00002fedc
}
\end{lstlisting}

In some rare cases, several different options for different vendors
may be specifed. In the folloging example 2 different values are
defined, depending on which vendor client will specify in \msg{SOLICIT} or
\msg{REQUEST} message. If client will only mention that it is interested in
any vendor specific into (i.e. did not sent \opt{vendor-spec info} option, but
only mentioned in in \opt{option request} option, it will receive
first vendor option defined (in the following example, that would be a
1234 and 0002fedc).

\begin{lstlisting}
# server.conf
log-level 8
log-mode precise
iface "eth1" {
 class {
   pool 2000::1-2000::ff
 }

 option vendor-spec 1234-0x00002fedc,5678-0x0002aaaa
}
\end{lstlisting}

\subsubsection{Example 12: Per client configuration}
\label{example-server-exceptions}
Usually all clients receive the same configuration options, e.g. all
clients will use the same DNS server. However, it is possible to
specify that particular clients should receive different options than
others. Following example set DNS server to 2000::1, domain
to example.com and vendor specific information for vendor 5678.
However, if requesting client has DUID 00:01:02:03:04:05:06:07:08, it
will receive different parameters (second.client.biz domain,
1234::5678:abcd as a DNS server and finally different vendor-specific
information). Also client with DUID 0x0001000044e8ef3400085404a324
will receives normal domain and DNS server, but different (vendor=2)
vendor specific information. See section \ref{feature-exceptions} for
background information. Since 0.8.0RC1, also addresses can be reserverd
in this way.

\begin{lstlisting}
# server.conf
iface "eth0" {

 class {
   pool 2000::1/64
 }

# common configuration options, provided for all clients
 option dns-server 2000::1
 option domain example.com
 option vendor-spec 5678-0x0002aaaa

  # special parameters for client with DUID 00:01:02:03:04:06
  client duid 0x000102030406
  {
        address 2001::123
        option domain second.client.biz
        option dns-server 1234::5678:abcd
        option vendor-spec 2-0x000122, 22-0x222222
  }

  # special address reserved for client with this remote-id option
  # (remote-id option may be added by relays)
  client remote-id 5-0x01020304
  {
        address 2000::0102:0304
        option domain our.special.remoteid.client.org
  }

  #  client link-local fe80::211:25ff:fe12:6688
  #  (link-local reservation is not supported yet. configure clients
  #  to use DUID-LL (link-local address based DUID) and then make
  #  DUID based reservations)
  #  {
  #      option domain link.local.detected.interop.test.com
  #  }

}
\end{lstlisting}

\subsubsection{Example 13: Prefix delegation}
\label{example-server-prefix}

Prefix delegation works quite similar to normal address granting.
Administrator defines pool and server provides prefixes from that
pool. Before using prefix delegation, please read section
\ref{feature-prefix}. Client configuration example is described in section
\ref{example-client-prefix}.

\begin{lstlisting}
# server.conf
log-mode precise

iface "eth0" {

 # the following lines instruct server to grant each client
 # prefix for this pool. For example, client might get
 # 2222:2222:2222:2222:2222:993f::/96
 pd-class {
        pd-pool 2222:2222:2222:2222:2222::/80
        pd-length 96
        T1 11111
        T2 22222
    }

}
\end{lstlisting}

\subsubsection{Example 14: Multiple prefixes}
\label{example-server-prefixes}
It is possible to define more than one pool, so each client will
receive several prefixes. It is necessary to define each pool with the
same length, i.e. it is not possible to mix different pool lengths.
See section \ref{feature-prefix} for prefix delegation background
information. Client configuration example is described in section
\ref{example-client-prefix}.

\begin{lstlisting}
# server.conf
log-mode precise

iface "eth0" {

 T1 1800
 T2 2700
 prefered-lifetime 3600
 valid-lifetime 7200

 # provide addresses from this pool
 class {
   pool 5000::/48
 }

 # the following lines instruct server to grant each client
 # 2 prefixes. For example, client might get
 # 2222:2222:2222:2222:2222:993f:6485::/96 and
 # 1111:1111:1111:1111:1111:993f:6485::/96
 pd-class {
        pd-pool 2222:2222:2222:2222:2222::/80
        pd-pool 1111:1111:1111:1111:1111::/80
        pd-length 96
        T1 11111
        T2 22222
    }

}
\end{lstlisting}

\subsubsection{Example 15: Inactive mode}
\label{example-server-inactivemode}
See sections \ref{example-client-inactivemode} and
\ref{feature-inactive-mode} for inactive mode explanation.
The same behavior has been added for server.

\begin{lstlisting}
#server.conf

log-level 8

inactive-mode

iface "eth0" {

 class {
   pool 2000::/64
 }
}
\end{lstlisting}

\subsubsection{Example 16: Leasequery}
A separate entity
called requestor can send queries regarding assigned addresses and
prefixes. Server can be configured to support such lease queries.
See section \ref{feature-leasequery} for detailed explanation.

\begin{lstlisting}
#server.conf

log-level 8

iface "eth0" {
 accept-leasequery

 class {
   pool 2000::/64
 }
}
\end{lstlisting}


\subsubsection{Example 17: Authentication}
\label{example-server-auth}
It is possible to configure server to require authentication. In this
example, HMAC-SHA-512 will be used as an authentication method.
Key Generation Nonce will have 64 bytes.

\begin{lstlisting}
# server.conf

auth-method digest-hmac-sha512
auth-key-len 64

iface eth0
{
  class
  {
    pool 2000::100-2000::10f
  }
}
\end{lstlisting}

\subsubsection{Example 18: Relay support with unknown interface-id}
\label{example-server-relay3}
To get more informations about relay configuration, see section \ref{feature-relays}.
In pervious examples (\ref{example-server-relay1},
\ref{example-server-relay2}) it was assumed that interface-id set by
relay is known. However, in some cases that is not true. If sysadmin
wants to accept relayed messages from any relay, there is a feature
called guess mode. It tries to match any relay defined in server.conf
instead of exactly checking interface-id value.

Since there is only one relay defined, it will be used, regardless of
the interface-id value (or even lack of thereof).

\begin{lstlisting}
# server.conf
guess-mode

iface relay1 {
  relay eth1
  interface-id 5020
  class {
    pool 2000::1-2000::ff
  }
  option dns-server 2000::100,2000::101
}
\end{lstlisting}


\subsubsection{Example 19: DS-Lite tunnel (AFTR)}
Server is able to provide Dual-Stack lite configuration for clients.
Both address and name based configurations are supported:

\begin{lstlisting}
iface "eth0" {
 class {
   pool 2001:db8::/64
 }

 option ds-lite 2001:db8:1::ffff
 option ds-lite sc.example.org
}
\end{lstlisting}

\subsubsection{Example 20: Custom options}
Server may be configured to also provide custom options to the
clients. See Section \ref{feature-custom-options} for details.

\begin{lstlisting}
iface "eth0" {
 class {
   pool 2001:db8::/64
 }
 option 145 - 01:02:a3:b4:c5:dd:ea
 option 146 address 2001:db8:1::dead:beef
 option 147 address-list 2001:db8:1::aaaa,2001:db8:1::bbbb
 option 148 string "secretlair.example.org"
}
\end{lstlisting}


\subsubsection{Example 21: Remote Autoconfiguration}
Server does support experimental extension called remote
autoconfiguration, as defined in \cite{draft-remote-autoconf}. See
Section \ref{feature-remote-autoconf} for details and configuration
examples.

\subsection{Relay configuration file}
\label{relay-conf}
Relay configuration is stored in \verb+relay.conf+ file in the
\verb+/etc/dibbler/+ directory (Linux systems) or in current directory
(Windows systems).

\subsubsection{Global scope}

Every option can be declared in global scope.
Config file consists of global options and one or more inteface
definitions. Note that reasonable minimum is 2 interfaces, as defining
only one would mean to resend messages on the same interface.

\subsubsection{Interface declaration}

Interface can be declared this way:
\begin{lstlisting}
iface interface-name
{
  interface options
}
\end{lstlisting}

or

\begin{lstlisting}
iface number
{
  interface options
}
\end{lstlisting}

where name\_of\_the\_interface denotes name of the interface and
number denotes it's number. It does not need to be enclosed in
single or double quotes (except windows cases, when interface name
contains spaces).

\subsubsection{Options}

Every option has a scope it can be used in, default value and
sometimes allowed range.

\begin{description}
 \item[log-level] -- (scope: global, type: integer, default: 7) Defines
            verbose level of the log messages. The valid range if
            from 1 (Emergency) to 8 (Debug). The higher the logging
            level is set, the more messages dibbler will print.
 \item[log-name] -- (scope: global, type: string, default: Client). Defines
            name, which should be used during logging.
 \item[log-mode] -- (scope: global, type: short, full or precise,
            default value: full) Defines logging mode. In the
            default, full mode, name, date and time in the h:m:s format
            will be printed. In short mode, only minutes and
            seconds will be printed (this mode is useful on
            terminals with limited width). Recently added precise
            mode logs information with seconds and microsecond
            precision. It is a useful for finding bottlenecks in
            the DHCPv6 autoconfiguration process.
\item[interface-id-order] -- (scope: global, type: before, after or omit,
        default: before) Defines placement of the
        interface-id option. Options can be placed in the \msg{RELAY-FORW}
        message is arbitrary order. This option has been specified to control
        that order. \opt{interface-id} option can be placed before or after
        \opt{relay-message} option. There is also possibility to instruct
        server to omit the \opt{interface-id} option altogether, but since
        this violates \cite{rfc3315}, it should not be used. In general, this
        configuration parameter is only useful when dealing with buggy relays,
        which can't handle all option orders properly. Consider this parameter
        a debugging feature. Note: similar parameter is available in the dibbler-server.

\item[client multicast] -- (scope: interface, type: boolean, default: false)
        This command instructs dibbler-relay to listen on this particular interface
        for client messages sent to multicast (ff02::1:2) address.
\item[client unicast] -- (scope: interface, type: address, default: not defined)
        This command instructs dibbler-relay to listen to messages sent to a specific
        unicast address. This feature is usually used to connect multiple relays
        together.
\item[server multicast] -- (scope: interface, type: boolean, default: false)
        This command instructs dibbler-relay to send messages (received on any interface)
        to the server multicast (ff05::1:3) address. Note that this is not the same
        multicast address as the server usually listens to (ff02::1:2). Server must be
        specifically configured to be able to receive relayed messages.
\item[server unicast] -- (scope: interface, type: address, default: none)
        This command instructs dibbler-relay to send message (received on any interface)
        to speficied unicast address. Server must be properly configured to to be able to
        receive unicast traffic. See \emph{unicast} command in the \ref{example-server-unicast}
        section.
\item[interface-id] -- (scope: interface, type: integer, default: none)
        This specifies identifier of a particular interface. It is used to generate
        \opt{interface-id} option, when relaying message to the server. This option
        is then used by the server to detect, which interface the message originates from.
        It is essential to have consistent interface-id defined on the relay side and
        server side. It is worth mentioning that interface-id should be specified on the
        interface, which is used to receive messages from the clients, not the
        one used to forward packets to server.
\item[guess-mode] -- (scope: global, type: boolean, default: no)
        Switches relay into so called guess-mode. Under normal operation, client sends
        messages, which are encapsulated and sent to the server. During this encapsulation
        relay appends \opt{interface-id} option and expects that server will use the same
        \opt{interface-id} option in its replies. Relay then uses those \opt{interface-id}
        values to detect, which the original request came from and sends reply to the same
        interface. Unfortunately, some servers does not sent \opt{interface-id} option.
        Normally in such case, dibbler-relay drops such server messages as there is no
        easy way to determine where such messages should be relayed to. However, when
        guess-mode is enabled, dibbler-relay tries to guess the destination interface.
        Luckily, it is often trivial to guess as there are usually 2 interfaces: one
        connected to server and second connected to the clients.
\end{description}

\subsection{Relay configuration examples}
\label{example-relay}

Relay configuration file is fairly simple. Relay forwards DHCPv6
messages between interfaces. Messages from client are encapsulated and
forwarded as RELAY\_FORW messages. Replies from server are received as
RELAY\_REPL message. After decapsulation, they are being sent back to
clients.

It is vital to inform server, where this relayed message was
received. DHCPv6 does this using interface-id option. This identifier
must be unique. Otherwise relays will get confused when they will
receive reply from server. Note that this id does not need to be
alligned with system interface id (ifindex). Think about it as
"ethernet segment identifier" if you are using Ethernet network or as
"bss identifier" if you are using 802.11 network.

If you are interested in additional examples, download source version
and look at \verb+*.conf+ files.

\subsubsection{Example 1: Simple}
\label{example-relay-1}
Let's assume this case: relay has 2 interfaces: eth0 and
eth1. Clients are located on the eth1 network. Relay should receive
data on that interface using well-known ALL\_DHCP\_RELAYS\_AND\_SERVER
multicast address (ff02::1:2). Note that all clients use multicast
addresses by default. Packets received on the eth1 should be
forwarded on the eth0 interface, using multicast address. See section
\ref{example-server-relay1} for corresponding server configuration.

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short
iface eth0 {
  server multicast yes
}
iface eth1 {
  client multicast yes
  interface-id 5020
}
\end{lstlisting}

\subsubsection{Example 2: Unicast/multicast}
It is possible to use unicast addresses instead/besides of default
multicast addresses. Following example allows message reception from
clients on the 2000::123 address. It is also possible to instruct
relay to send encapulated messages to the server using unicast
addresess. This feature is configured in the next section
(\ref{example-relay-multiple}).

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short
iface eth0 {
  server multicast yes
}
iface eth1 {
  client multicast yes
  client unicast 2000::123
  interface-id 5020
}
\end{lstlisting}

\subsubsection{Example 3: Multiple interfaces}
\label{example-relay-multiple}
Here is another example. This time messages should be forwarded from
eth1 and eth3 to the eth0 interface (using multicast) and to the eth2
interface (using server's global address 2000::546). Also clients must
use multicasts (the default approach):

\begin{lstlisting}
# relay.conf
iface eth0 {
  server multicast yes
}
iface eth2 {
  server unicast 2000::456
}
iface eth1 {
  client multicast yes
  interface-id 1000
}
iface eth3 {
  client multicast yes
  interface-id 1001
}
\end{lstlisting}

\subsubsection{Example 4: 2 relays}
\label{example-relay-cascade}
Those two configuration files correspond to the ,,2 relays'' example
provided in section \ref{example-server-relay2}. See section
\ref{feature-relays} for detailed exmplanations.

\begin{lstlisting}
# relay.conf - relay 1
log-level 8
log-mode full

# messages will be forwarded on this interface using multicast
iface eth2 {
   server multicast yes    // relay messages on this interface to ff05::1:3
 # server unicast 6000::10 // relay messages on this interface to this global address
}

iface eth1 {
#  client multicast yes    // bind ff02::1:2
  client unicast 6011::1   // bind this address
  interface-id 6011
}
\end{lstlisting}

\begin{lstlisting}
# relay.conf - relay 2
iface eth0 {
#   server multicast yes  // relay messages on this interface to ff05::1:3
  server unicast 6011::1  // relay messages on this interface to this global address
}

# client can send messages to multicast
# (or specific link-local addr) on this link
iface eth1 {
  client multicast yes    // bind ff02::1:2
# client unicast 6021::1  // bind this address
  interface-id 6021
}
\end{lstlisting}

\subsubsection{Example 5: Guess-mode}
In the 0.6.0 release, a new feature called guess-mode has been
added. When client sends some data and relay forwards it to the
server, it always adds interface-id option to specify, which link
the data has been originally received on. Server, when responding to
such request, should include the same interface-id option in the
reply. However, in some poor implementations, server fails to do
that. When relay receives such poorly formed response from the server,
it can't decide which interface should be used to relay this
message.

Normally such packets are dropped. However, it is possible to switch
relay into a guess-mode. It tries to find any suitable interface,
which it can forward data on. It is not very reliable, but sometimes
it is better than dropping the message altogether.

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short
guess-mode

iface eth0 {
  server multicast yes
}
iface eth1 {
  client multicast yes
  interface-id 5020
}
\end{lstlisting}

\subsubsection{Example 6: Relaying to multicast}
During normal operation, relay sends forwarded messages to a
\emph{All\_DHCP\_Servers} (FF05::1:3) multicast address.

Although author does not consider this an elegant solution, it is also
possible to instruct relay to forward message to a \emph{All\_DHCP\_Relay\_Agents\_and\_Servers}
(ff02::1:2) multicast address. That is quite convenient when there are several
relays connected in a cascade way (server -- relay1 -- relay2 -- clients).

For details regarding DHCPv6-related multicast addresses and relay operation, see \cite{rfc3315}.

To achieve this behavior, \emph{server unicast} can be used. Note that
name of such parameter is a bit misleading (``server unicast'' used to specify
multicast address). That parameter should be rather called ``destination address'',
but to maintain backward compatibility, it has its current name.

\begin{lstlisting}
# relay.conf
log-level 8
log-mode short

iface eth0 {
  server unicast ff02::1:2
}
iface eth1 {
  client multicast yes
  interface-id 5020
}
\end{lstlisting}

\subsection{Requestor configuration}
Requestor (entity used for leasequery) does not use configuration
files. All parameters are specified by command-line switches. See
section \ref{feature-leasequery} for details.
